---
title: "HB_dynatop"
format: html
editor_options: 
  chunk_output_type: console
---

Model runs, but there is something weird going on. The result is always extremely low. I believe units are correct.

```{r}
library(pacman)
p_load(terra, dynatop, dynatopGIS, xts, whitebox, tidyverse,
       lubridate,
       sf, # working with st and sf objects
       tidyterra, # do tidy ops on terra objects
       sfnetworks # create start and end nodes for channel network
       )
setwd("./hb_data")
```

```{r}
# Try reading my own version of dynatop
devtools::load_all(file.path(".","packages", "dynatop-master")) ## attach the revised package

```


# GIS PREP
```{r GIS files}
dem_path <- file.path(".", "unprocessed",
                      "1m hydro enforced dem","hydem1mlpns.tif")
dem <- rast(dem_path) # load the dem as a raster layer
plot(dem)

# crop DEM to a rough rectangle
ybounds <- c(4870350,4871350)
xbounds <- c(281350, 282150)
crop1 <- crop(dem, ext(c(xbounds, ybounds)))
plot(crop1)
crop_dem_path <- file.path(".", "unprocessed",
                      "1m hydro enforced dem","hydem1mlpns_cropped.tif")
writeRaster(crop1, crop_dem_path, overwrite = TRUE)
####delineate watershed and keep watershed boundary

w3_breached <- "./unprocessed/whitebox_outputs/1mdem_breach.tif"
wbt_breach_depressions_least_cost(
  dem = crop_dem_path,
  output = w3_breached,
  dist = 1,
  fill = TRUE)

w3_filled <- "./unprocessed/whitebox_outputs/1mdem_fill.tif"
wbt_fill_depressions_wang_and_liu(
  dem = w3_breached,
  output = w3_filled
)
#calculate flow accumulation and direction
w3_flowacc <- "./unprocessed/whitebox_outputs/1mdem_w3_flowacc.tif"
wbt_d8_flow_accumulation(input = w3_filled,
                         output = w3_flowacc)
w3_d8pt <- "./unprocessed/whitebox_outputs/1mdem_w3_d8pt.tif"
wbt_d8_pointer(dem = w3_filled,
               output = w3_d8pt)

#delineate streams
w3_streams <- "./unprocessed/whitebox_outputs/w3_streams.tif"
wbt_extract_streams(flow_accum = w3_flowacc,
                    output = w3_streams,
                    threshold = 8000)
plot(rast(w3_streams))

# convert streams to shapefile
w3_streams_vect <- "./unprocessed/whitebox_outputs/w3_streams_vect.shp"
wbt_raster_streams_to_vector(streams = w3_streams,
                             d8_pntr = w3_d8pt,
                             output = w3_streams_vect)

# get pour point ready
pour_point_path <- file.path(".", 
                             "unprocessed",
                             "1m hydro enforced dem", "w3_pour_snap.shp")
plot(vect(pour_point_path), add = TRUE)

w3_pour_snap <- "./unprocessed/whitebox_outputs/w3_pour_snap.shp"
wbt_jenson_snap_pour_points(pour_pts = pour_point_path,
                            streams = w3_streams,
                            output = w3_pour_snap,
                            snap_dist = 10)
w3_pour_snap_read <- vect(w3_pour_snap)
plot(rast(w3_streams), 
     xlim = c(281400, 282000),
     ylim = c(4870400, 4870800))
points(w3_pour_snap_read, pch = 1)


w3_shed <- "./unprocessed/whitebox_outputs/w3_shed.tif"
wbt_watershed(d8_pntr = w3_d8pt,
              pour_pts = w3_pour_snap,
              output = w3_shed)

plot(rast(w3_shed))
#convert raster of watershed area to vector for final mapping
w3_outline <- as.polygons(rast(w3_shed), extent=FALSE)
plot(w3_outline)

plot(w3_outline)

channel <- vect(w3_streams_vect) %>% 
  crop(w3_outline)


# crop and mask DEM by watershed outline
dem_crop_mask <- rast(w3_filled) %>% 
  crop(w3_outline) %>% 
  mask(w3_outline)

#create watershed area, just use area of DEM
outline <- (dem_crop_mask * 0) + 1
# extend both rasters by 1 pixel
outline_extended <- extend(outline,1)
dem_final <- extend(dem_crop_mask,1)


writeRaster(outline_extended,file.path(".","processed","w3.tif"),overwrite=TRUE)
writeRaster(dem_final,file.path(".","processed","dem.tif"),overwrite=TRUE)
```
```{r convert channel}
plot(channel)
channel <- channel %>% 
  mutate(ID = seq(1, n(), 1))
# 2. Break multipart lines into individual components
# This converts MULTILINESTRING to LINESTRING
single_parts <- st_cast(as_sf(channel), "LINESTRING")

# 3. View the new, broken-down data
print(single_parts)
plot(single_parts)
coords <- st_coordinates(single_parts)


net <- as_sfnetwork(single_parts, directed = TRUE)

# Extract nodes
nodes <- st_as_sf(net, "nodes")

# Extract edges
edges <- st_as_sf(net, "edges")

edges$length_m <- st_length(edges)
test <- 
  edges %>% 
  select(from, to, geometry, length_m) %>% 
  mutate("Identifier" = dplyr::row_number()) %>% 
  vect() #%>%

crs(test) <- crs(dem_final)

conv_channel <- convert_channel(
    test,
    property_names = c(name = "Identifier", 
                       length = "length_m", startNode = "from",
                       endNode= "to"),
    default_width = 1,
    default_slope = 0.001
)

writeVector(conv_channel,file.path(".","processed","channel.shp"),overwrite=TRUE)

# name is different than the start and end nodes!
conv_channel %>% 
  as.data.frame() %>% 
  tibble()
```

```{r}
# Find the outlet ID right after you create your 'edges' object
edges_df <- st_drop_geometry(edges)
outlet_edge <- edges_df %>% filter(!(to %in% from))
outlet_id <- outlet_edge$ID

# Update your output definition using the correct outlet ID
output_defn <- data.frame(name = "W3",
                          id = outlet_id,
                          flux = "q_sf",
                          scale = 1)


## create a vector of the uid's of the HRUs
hid <- sapply(mdl$hru,function(h){h$id})
## create a vector of the channel names with as storaged in the class variabel
cname <- sapply(mdl$hru,function(h){h$class$name})
idx <- match(gauges$chn_identi, cname) ## index of HRU's corresponding to gauges
## create as a data.frame
gauge_output <- data.frame(name = gauges$Site.Name,
                           id = as.integer(hid[idx]),
                           flux = "q_sf",
                           scale = 1)
saveRDS(gauge_output,file.path(".","dyna","gauge_output_defn.rds")) ## save the output for later use
gauge_output
```

Create in the same resolution rasters of:
- watershed shape
- DEM
Shapefile of:
- pour point/gauge location
- channel



# OBS PREP
```{r}

flow <- as.xts(read.zoo(file.path(".", "unprocessed","HBEF_DailyStreamflow_1956-2024.csv"),
                        header=TRUE,sep=",",drop=FALSE)) 
head(flow)
flow_w3 <- flow[flow$WS == 3, ]
subset_flow <- flow_w3["2003-03-01/2005-03-01"] 
subset_flow[,2] <- subset_flow[,2] #* 0.003404861
```

```{r precip and discharge}
# discharge
# mm/day
subset_flow <- 
  read_csv(file.path(".", 
                   "unprocessed",
                   "HBEF_DailyStreamflow_1956-2024.csv")) %>% 
  filter(DATE >= "2003-03-01",
         DATE <= "2005-03-01",
         WS == 3) %>% 
  mutate(Streamflow = Streamflow / 1000) %>% 
  select(DATE, Streamflow) %>% 
  as.xts()
```
```{r PRECIP}
# precipitation
# m/day
subset_precip <- read_csv(file.path(".", 
                   "unprocessed",
                   "dailyWatershedPrecip1956-2025.csv")) %>% 
  filter(DATE >= "2003-03-01",
         DATE <= "2005-03-01",
         watershed == "W3") %>% 
  mutate(precip = Precip / 1000) %>% 
  select(DATE, precip) %>% 
  as.xts()
```

```{r PET}
# PET
# m/day




# --- Inputs -------------------------------------------------------------

temps <- read_csv("unprocessed/HBEF_air_temp_daily.csv") %>% 
  filter(date >= "2003-03-01",
         date <= "2005-03-01",
         STA == "STA6") %>% 
  select(date, MAX, MIN)
write_csv(temps, "unprocessed/air_temp_sta6.csv")

# example data frame
met <- data.frame(
  date = temps$date,
  tmin = temps$MIN,   # °C
  tmax = temps$MAX   # °C
)

lat <- 45   # latitude in decimal degrees (e.g., Blacksburg VA)

# --- Hargreaves PET -----------------------------------------------------

hargreaves_pet <- function(date, tmin, tmax, lat){

  J <- yday(date)                 # day of year
  phi <- lat * pi/180             # latitude radians
  
  # extraterrestrial radiation (Ra)
  dr  <- 1 + 0.033*cos(2*pi*J/365)
  delta <- 0.409*sin(2*pi*J/365 - 1.39)
  ws <- acos(-tan(phi)*tan(delta))

  Gsc <- 0.0820   # MJ m-2 min-1
  
  Ra <- (24*60/pi) * Gsc * dr *
        (ws*sin(phi)*sin(delta) +
         cos(phi)*cos(delta)*sin(ws))

  tmean <- (tmin + tmax)/2

  PET <- 0.0023 * Ra *
         (tmean + 17.8) *
         sqrt(pmax(tmax - tmin, 0))

  return(PET)   # mm/day
}

# --- Calculate PET ------------------------------------------------------

met$PET_mm_day <- hargreaves_pet(
  met$date,
  met$tmin,
  met$tmax,
  lat
)

met$PET_01 <- (met$PET_mm - min(met$PET_mm, na.rm=TRUE)) /
              (max(met$PET_mm, na.rm=TRUE) -
               min(met$PET_mm, na.rm=TRUE))

met %>% 
  #mutate(pet = 0) %>% # PET in m
  select(date, PET_mm_day) %>% 
  mutate(PET_m_day = PET_mm_day / 1000) %>% 
  rename(pet = PET_m_day) %>% 
  select(date, pet) %>% 
  write_csv("unprocessed/pet.csv")

pet <- read_csv(file.path(".", 
                   "unprocessed",
                   "pet.csv")) %>% 
  as.xts()
 
```

```{r combine-datasets}
obs <- merge(subset_flow,
             subset_precip,
             pet,
             all=FALSE) ## merge by time stamp
#obs <- obs[ , colSums(is.na(obs)) == 0]               # Remove columns with at least one NA

# THE FIX: Force the index to be POSIXct so dynatop calculates dt = 86400 seconds
index(obs) <- as.POSIXct(as.character(index(obs)), tz = "UTC")

# Save it
saveRDS(obs, file.path(".", "processed", "obs.rds"))

```

# CREATE DYNATOP OBJECT
```{r}
ctch <- dynatopGIS$new(file.path(".", "dynaGIS"))
# add catchment, DEM, channel
ctch$add_catchment( file.path(".","processed","w3.tif") )
ctch$add_dem(file.path(".","processed","dem.tif"))
ctch$add_channel(file.path(".","processed","channel.shp"))

ctch$get_layer()


ctch$plot_layer("dem", add_channel=TRUE)
```

```{r fill-sinks}
ctch$sink_fill()

terra::plot( ctch$get_layer('filled_dem') - ctch$get_layer('dem'),
             main="Changes to height")
```

```{r compute-stuff}
# bands
ctch$compute_band()
ctch$plot_layer("band")
# twi
ctch$compute_properties()
ctch$plot_layer('atb')
# flow lengths
ctch$compute_flow_lengths(flow_routing="shortest")
ctch$plot_layer("shortest_flow_length")
```

```{r classifying and cutting}
## simple equally spaced cuts
ctch$classify("atb_20","atb",20)
ctch$plot_layer("atb_20")

## use the quantile function of the raster package to compute breaks
## note we need to include the min and max
brk <- global(ctch$get_layer("atb"), fun=quantile,probs = seq(0, 1, length = 21),na.rm=T)
ctch$classify("atb_20_equal","atb",brk)
ctch$plot_layer("atb_20_equal")

ctch$combine_classes("atb_20_subcatch",c("atb_20","catchment"))
ctch$plot_layer("atb_20_subcatch")

ctch$combine_classes("atb_20_subcatch_band",c("atb_20_subcatch","band"))
ctch$plot_layer("atb_20_subcatch_band")


# Look at classes
tmp <- ctch$get_method("atb")
tmp$type
#> [1] "combination"
head(tmp$groups)

```

```{r create the model}
ctch$create_model(file.path("./dynaGIS","atb_20"), # name of new model
                  "atb_20", # classification to base model on
                  sf_opt="kin",
                  sz_opt="exp",
                  )

## load the model
mdl <- readRDS( file.path(".","dynaGIS","atb_20.rds")) ## read the model in
names(mdl)
```


# RUN MODEL
```{r run model and plot output}
# Load previously created data
output_defn <- data.frame(name = "W3",
                      id = 1,
                      flux = "q_sf",
                      scale = 1)
mdl <- readRDS(file.path(".","dynaGIS","atb_20.rds"))
obs <- readRDS(file.path(".","processed","obs.rds"))
# initialize dynatop object
hru <- mdl$hru
ctch_mdl <- dynatop$new(hru)

ctch_mdl$add_data(obs)
## determining the initial recharge rate ad setting it in the model
# I think the same number is the watershed area...
#r0 <- as.numeric(obs[1, "Streamflow"])#/ 86400
#r0 <- as.numeric(obs$Streamflow[1])
 r0 <- as.numeric(obs$Discharge_m3_day[1])/ sum( sapply( hru, function(x){x$properties["area"]} ) )
hru <- lapply(hru,function(h){h$initialisation["r_uz_sz_0"] <- r0;h})

## the map input is only needed for visualisation
ctch_mdl <- dynatop$new( hru, map = mdl$map
  )$add_data( obs
  )$initialise(
  )$sim(output_defn
  )


# Gauge Flows
sim_gauge <- ctch_mdl$get_output() ## extract the flow as an xts object
head(sim_gauge)
plot( merge(obs$Discharge_m3_day,sim_gauge),main="Simulated and observed flows",
     legend.loc='topright')


```

```{r}
# RUN MODEL
mdl <- readRDS(file.path(".","dynaGIS","atb_20.rds"))
obs <- readRDS(file.path(".","processed","obs.rds"))
hru <- mdl$hru

#hru <- lapply(hru, function(h){ h$uz$parameters["t_d"] <- 8 * 60 * 60; h})
# 1. Use the outlet_id you generated in your GIS prep chunk
output_defn <- data.frame(name = "W3",
                          id = 1, # <-- USE YOUR VARIABLE HERE
                          flux = "q_sf",
                          scale = 1)


r0 <- as.numeric(obs$Discharge_m3_s[1])
hru <- lapply(hru, function(h) { h$initialisation["r_uz_sz_0"] <- r0; h })

# 3. Run the model
ctch_mdl <- dynatop$new(hru, map = mdl$map)$
  add_data(obs)$
  initialise()$
  sim(output_defn,
      sub_step = 900)

# 4. Extract and scale flow
sim_rate <- ctch_mdl$get_output() 
sim_vol_m3_day <- sim_rate * 86400 

catchment_area <- sum(sapply(hru, function(x) x$properties["area"]))
sim_specific_m_day <- sim_vol_m3_day / catchment_area

plot(ctch_mdl$get_output(), type = 'p')
# 5. Plot
plot(merge(obs$Discharge_m3_s, sim_rate), 
     main="Simulated vs Observed Specific Discharge (m/day)",
     legend.loc='topleft')
```

```{r map}
# 1. Load the model object
mdl <- readRDS(file.path(".", "dynaGIS", "atb_20.rds"))

# 2. Read the spatial map of HRUs 
# (mdl$map stores the file path to the HRU tif generated by dynatopGIS)
hru_map <- rast(mdl$map) 

# Rename the raster layer so we can call it easily in ggplot
names(hru_map) <- "hru_id"

# 3. Convert the raster to polygons to calculate the center of each HRU
hru_poly <- as.polygons(hru_map)

# Calculate the centroids (center points) of each polygon for the text labels
hru_centroids <- centroids(hru_poly)

# Extract coordinates into a standard dataframe for ggplot
centroid_coords <- crds(hru_centroids, df = TRUE)
centroid_coords$hru_id <- hru_centroids$hru_id

# 4. Plot using tidyterra and ggplot2
ggplot() +
  # Plot the raster base map (using factor for discrete colors)
  geom_spatraster(data = as.factor(hru_map)) +
  # Overlay the polygon boundaries
  geom_spatvector(data = hru_poly, fill = NA, color = "black", linewidth = 0.3) +
  # Add the ID labels at the center of each HRU
  geom_text(data = centroid_coords, aes(x = x, y = y, label = hru_id), 
            color = "white", fontface = "bold", size = 4) +
  # Use a nice color palette and remove the bulky legend
  scale_fill_viridis_d(option = "turbo", na.value = "transparent", guide = "none") +
  theme_minimal() +
  labs(title = "Hubbard Brook HRU & Channel ID Map",
       subtitle = "Hillslopes (1-20) and Channels (21+)",
       x = "Longitude", y = "Latitude")
```

```{r Mass Balance}
mb <- ctch_mdl$get_mass_errors()
head(mb)
tail(mb)
plot( mb$outflow, main="Mass balance error",ylab="m^3")
plot( mb$e_t, main="Mass balance error",ylab="m^3")

plot( mb$p/catchment_area , main="Mass balance error",ylab="m^3")

plot( mb$initial_state , main="Mass balance error",ylab="m^3")

mdl$output_flux

ctch_mdl$plot_output()
```

```{r}
## extract the current states as a data.frame
current_state <- ctch_mdl$get_states()
## rows correspond to the Hill-slope HRU ID, and the storage's in each of the four zones
head(current_state)

ctch_mdl$plot_state("s_sf")
ctch_mdl$plot_state("s_rz")
ctch_mdl$plot_state("s_uz")
ctch_mdl$plot_state("s_sz")
```

```{r changing parameters}

chn_id <- do.call(c, lapply( hru, function(h){ if(!is.na(h$class$startNode)){ h$id } }))

out_hyd <-  data.frame( name = paste0("hyd_",chn_id),
                       id = chn_id,
                       flux = "v_sf_rz", scale =-1/900) # scale is negative since flux usually downwards, 1/900 is 1/timestep

out_hyd <-  data.frame( name = paste0("hyd_",chn_id),
                       id = chn_id,
                       flux = "v_sf_rz", scale =-1/900)
output_defn <- out_hyd[2,]
## function to compute the relevent part so the output definition for each potential source HRU
fout <- function(h){
    ## test if it is a channel HRU - in which case return NULL
    if( !is.na(h$class$startNode) ){ return(NULL) }
        
    ## index of flow directiuons going to other channels
    idx <- h$sf_flow_direction$id %in% chn_id
    if(!any(idx)){ return(NULL) } ## quit if no links!
    ## create partial output defn
    data.frame(
        name = paste0("hyd_",h$sf_flow_direction$id[idx]), ## which inflow does it alter
        id = h$id, ## flux comes from this HRU
        flux = "q_sf", # surface flux out of this HRU is input to the next
        scale = h$sf_flow_direction$fraction[idx] # scale is fraction of flow (since taking it away)
    )
}
## row bind to give a data.frame
tmp <- do.call( rbind, lapply( hru, fout) )

## row bind onto the intial data frame
out_hyd <- rbind( out_hyd, tmp )

## sort this, not needed but easier to look at
out_hyd <- out_hyd[ order(out_hyd$name), ]

ctch_mdl <- dynatop$new(hru)$add_data(obs)$initialise()$sim(out_hyd)
plot( ctch_mdl$get_output(), legend.loc='topright')
```

# MONTE CARLO PARAMETERS
```{r change parameters function}
## h is the hru, p is a named vector of parameter values
fparChange <- function(h,p){
    ## change "m" transmissivty decay parameter depending upon atb class
    
    h$sz$parameters["m"] <- p["m1"]
       
    ## change unsaturated zone time constant regarless of type
    h$uz$parameters["t_d"] <- p["Td"]
    ## change surface velocity depending upon if it is a channel
    if( is.na(h$class$startNode) ){
        ## hillslope
        h$sf$parameters["c_sf"] <- p["csf"]
    }else{
        ## channel
        h$sf$parameters["c_sf"] <- p["vch"]
    }
    return(h)
}
```

```{r randomly simulate parameters}
## start by loading the model for reference
nsim <- 100
sample <- cbind(
  sid = 1:nsim, ## the simulation identifier number
  Td = runif(nsim,0.1,120),
  m1 = runif(nsim,0.005,0.06),
  m2 = runif(nsim,0.005,0.06),
  csf = runif(nsim,0.05,0.5),
  vch = runif(nsim,0.2,1.4))
## save the random sample
#saveRDS(sample,file.path(".","dynaGIS","sample.rds"))

```

```{r}
## normally this function would be entered into its own file.
## Here to show it we enter it then add it to a file
## inputs are i the index of a row of sample and the model - which is altered by the function
## It expects the sample, pattern, obs and q0 to be present in the workspace fMC is called from
## it does not need its own copy of these since they are not changed
fMC <- function(i,sample,mdl,obs,q0,fP){
#i <- 1
  x <- sample[i,] # the values to substitute in
  ## put sample values into model
  hru <- lapply( mdl$hru, fparChange, p=x) 
  ## put intial conditions into the model
  hru <- lapply(hru,function(h,q0){h$initialisation["r_uz_sz_0"] <- q0; h},q0=q0)
    ## simulate the model
  dt <- dynatop::dynatop$new(hru)$add_data(obs)$initialise()$sim(output_defn)
  ## get mass balance
  mb <- dt$get_mass_errors()

  plot( mb$outflow, main="Mass balance error",ylab="m^3")

  ##ch_in <- dt$get_channel_inflow()
  gf <- dt$get_output()

  ## ns = nash-sutcliffe #q_sf_0
  ns <- 1 - sum((obs$Discharge_m3_s-gf)^2,na.rm=TRUE) /
    sum((obs$Discharge_m3_s - mean(obs$Discharge_m3_s,na.rm=TRUE))^2,na.rm=TRUE)

  return(ns)
}

x <- sample[2,]
hru <- lapply( mdl$hru, fparChange, p=x) 
# 3. Run the model
ctch_mdl <- dynatop$new(hru)$
  add_data(obs)$
  initialise()$
  sim(output_defn,
      sub_step = 900)
  plot(merge(ctch_mdl$get_output(), obs$Discharge_m3_s),  main="Mass balance error",ylab="m^3",
        legend.loc='topright')

1 - sum((obs$Discharge_m3_s-ctch_mdl$get_output())^2,na.rm=TRUE) /sum((obs$Discharge_m3_s - mean(obs$Discharge_m3_s,na.rm=TRUE))^2,na.rm=TRUE)

```

```{r}
rm(list=ls())

output_defn <- data.frame(name = "W3",
                          id = 1,
                          flux = "q_sf",
                          scale = 1)

# load model and obs
mdl <- readRDS(file.path(".","dynaGIS","atb_20.rds"))
obs <- readRDS(file.path(".","processed","obs.rds"))

catchment_area <- sum(sapply(mdl$hru, function(x) x$properties["area"]))
q0 <- as.numeric(obs$Discharge_m3_s[1]) / catchment_area
## this wraps the fMC function to catch any errors and stop then killing the simulation 


fout <- function(i,sample,mdl,obs,q0,fP){
    tryCatch({fMC(i,sample,mdl,obs,q0,fP)},
             error=function(e){e$message}) ## if an error return the message
}

fout(2, sample, mdl, obs, q0, fparChange)
fMC(1,sample,mdl,obs,q0,fparChange)
## use lapply to iterate over the parameter sets - this will run on a single R process
#out <- lapply(seq_len(nrow(sample)), fout)
out <- lapply(seq_len(nrow(sample)), fout, sample = sample, mdl = mdl, obs = obs, q0 = q0, fP = fparChange)
## on linux/mac use mclapply to run across multiple cores on the same machine
## alter mc.cores to match the machine being used
## out <- mclapply(1:6,fMC,sample=sample,mdl=mdl,obs=obs,q0=q0,fP=fparChange,mc.cores=6,mc.preschedule = FALSE)

## on windows/linux/mac you could use parLapply but the set up is more complex
cl <- makePSOCKcluster(6) # alter number to be number of cores on machine you want to use
out2 <- parLapply(cl,1:6,fMC,sample=sample,mdl=mdl,obs=obs,q0=q0,fP=fparChange)
stopCluster(cl)
## The `future` package offer an alternative.


## save out after naming with the simulation number
names(out) <- paste0("sid_",sample[,"sid"])
saveRDS(out,"output.rds")
```

Notes from 4-9-26, close to figuring it out I think
Modifying input parameters is not really changing anything, maybe there is something else critically wrong with model. There is something weird with units for sure. 

Unclear if simulations are actually using new parameter set, it seems I can't check it once it has run.

I think that the weirdness is not from the parameters, or at least none of the ones I have tried changing so far.

