---
title: "HB_dynatop7"
format: html
editor_options: 
  chunk_output_type: console
---
Difference from dynatop6:
- making parameterization as simple as possible
- saving final formatted model object, read in on each run. This gurantees it is the same thing
- test to make sure switching parameters works as intended

# ONLY RUN ONCE- CREATING SETUP FOR RUNNING MODEL
```{r setup}
# Load packages
library(pacman)
p_load(terra, xts, whitebox, tidyverse,
       lubridate,
       dynatopGIS,
       sf, # working with st and sf objects
       tidyterra, # do tidy ops on terra objects
       sfnetworks, # create start and end nodes for channel network
       lwgeom,
       lhs, DEoptim, # maybe obsolete since I am using GLUE
       parallel,
       readr
       )

# Load modified version of dynatop, which includes snow routine
devtools::load_all(file.path("/Users/johnmorgan/Documents/HBTopModel/revised_package/dynatop")) ## attach the revised package

crs1 <- "PROJCRS[\"NAD83 / UTM zone 19N\",\n    BASEGEOGCRS[\"NAD83\",\n        DATUM[\"North American Datum 1983\",\n            ELLIPSOID[\"GRS 1980\",6378137,298.257222101,\n                LENGTHUNIT[\"metre\",1]]],\n        PRIMEM[\"Greenwich\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433]],\n        ID[\"EPSG\",4269]],\n    CONVERSION[\"UTM zone 19N\",\n        METHOD[\"Transverse Mercator\",\n            ID[\"EPSG\",9807]],\n        PARAMETER[\"Latitude of natural origin\",0,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8801]],\n        PARAMETER[\"Longitude of natural origin\",-69,\n            ANGLEUNIT[\"degree\",0.0174532925199433],\n            ID[\"EPSG\",8802]],\n        PARAMETER[\"Scale factor at natural origin\",0.9996,\n            SCALEUNIT[\"unity\",1],\n            ID[\"EPSG\",8805]],\n        PARAMETER[\"False easting\",500000,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8806]],\n        PARAMETER[\"False northing\",0,\n            LENGTHUNIT[\"metre\",1],\n            ID[\"EPSG\",8807]]],\n    CS[Cartesian,2],\n        AXIS[\"(E)\",east,\n            ORDER[1],\n            LENGTHUNIT[\"metre\",1]],\n        AXIS[\"(N)\",north,\n            ORDER[2],\n            LENGTHUNIT[\"metre\",1]],\n    USAGE[\n        SCOPE[\"Engineering survey, topographic mapping.\"],\n        AREA[\"North America - between 72°W and 66°W - onshore and offshore. Canada - Labrador; New Brunswick; Nova Scotia; Nunavut; Quebec. Puerto Rico. United States (USA) - Connecticut; Maine; Massachusetts; New Hampshire; New York (Long Island); Rhode Island; Vermont.\"],\n        BBOX[14.92,-72,84,-66]],\n    ID[\"EPSG\",26919]]"
```

## GIS PREP
```{r GIS files}
target_folder <- "./7th_attempt/unprocessed/"
dem_path <- file.path(".", "hb_data", "unprocessed",
                      "1m hydro enforced dem","hydem1mlpns.tif")
dem <- rast(dem_path) # load the dem as a raster layer

# crop DEM to a rough rectangle
ybounds <- c(4870350,4871350)
xbounds <- c(281350, 282150)
crop1 <- crop(dem, ext(c(xbounds, ybounds)))
plot(crop1)

dem3 <- terra::aggregate(crop1, 5, fun = "mean")
#save raster, because whitebox wants it is a files location instead of an object in R
crop_3mdem_path <- paste0(target_folder, "3mdem.tif")
writeRaster(dem3, crop_3mdem_path, overwrite = TRUE)

####delineate watershed and keep watershed boundary

w3_breached <- paste0(target_folder,"3mdem_breach.tif")
wbt_breach_depressions_least_cost(
  dem = crop_3mdem_path,
  output = w3_breached,
  dist = 5,
  fill = TRUE)

w3_filled <- paste0(target_folder,"3mdem_fill.tif")
wbt_fill_depressions_wang_and_liu(
  dem = w3_breached,
  output = w3_filled
)
#calculate flow accumulation and direction
w3_flowacc <- paste0(target_folder,"3mdem_flowacc.tif")
wbt_d8_flow_accumulation(input = w3_filled,
                         output = w3_flowacc)
w3_d8pt <- paste0(target_folder,"3mdem_d8pt.tif")
wbt_d8_pointer(dem = w3_filled,
               output = w3_d8pt)

#delineate streams
w3_streams <- paste0(target_folder,"streams.tif")
wbt_extract_streams(flow_accum = w3_flowacc,
                    output = w3_streams,
                    threshold = 200)
plot(rast(w3_streams))

# convert streams to shapefile
w3_streams_vect <- paste0(target_folder,"streams_vect.shp")
wbt_raster_streams_to_vector(streams = w3_streams,
                             d8_pntr = w3_d8pt,
                             output = w3_streams_vect)
plot(vect(w3_streams_vect))

# get pour point ready
pour_point_path <- file.path(".", "hb_data",
                             "unprocessed",
                             "1m hydro enforced dem", "w3_pour_snap.shp")
plot(vect(pour_point_path), add = TRUE)

w3_pour_snap <- paste0(target_folder,"pour.shp")
wbt_jenson_snap_pour_points(pour_pts = pour_point_path,
                            streams = w3_streams,
                            output = w3_pour_snap,
                            snap_dist = 10)
w3_pour_snap_read <- vect(w3_pour_snap)
plot(rast(w3_streams), 
     xlim = c(281400, 282000),
     ylim = c(4870400, 4870800))
points(w3_pour_snap_read, pch = 1)


w3_shed <- paste0(target_folder,"shed.tif")
wbt_watershed(d8_pntr = w3_d8pt,
              pour_pts = w3_pour_snap,
              output = w3_shed)

plot(rast(w3_shed))
#convert raster of watershed area to vector for final mapping
w3_outline <- as.polygons(rast(w3_shed), extent=FALSE)
plot(w3_outline)

plot(w3_outline)

channel <- vect(w3_streams_vect, crs = crs1) %>% 
  crop(w3_outline)


# crop and mask DEM by watershed outline
dem_crop_mask <- rast(w3_filled) %>% 
  crop(w3_outline) %>% 
  mask(w3_outline)

#create watershed area, just use area of DEM
outline <- (dem_crop_mask * 0) + 1
# extend both rasters by 1 pixel
outline_extended <- extend(outline,1)
dem_final <- extend(dem_crop_mask,1)

# STIC locations
all_24 <- read_csv("/Users/johnmorgan/Documents/VT Research/HBTopModel/STIC_Synthesis/all_24.csv")

stic_locs <- all_24 %>%
  filter(wshed == "W3") %>%
  select(number, lat, long) %>%
  distinct() %>%
  vect(geom = c("long", "lat"), crs = "EPSG:4326") %>%
  project("EPSG:26919")
# save STIC locations, un-snapped
stic_locs_not_snapped <- paste0(target_folder,"not_snapped_stics.shp")
writeVector(stic_locs, stic_locs_not_snapped, overwrite = TRUE)
# snap STIC sensors to channel network
stic_locs_snapped <- paste0(target_folder,"snapped_stics.shp")

wbt_jenson_snap_pour_points(pour_pts = stic_locs_not_snapped,
                            streams = w3_streams,
                            output = stic_locs_snapped,
                            snap_dist = 30)
ggplot() +
  # Plot the raster base map (using factor for discrete colors)
  geom_spatvector(data = w3_outline) +
    geom_spatvector(data = channel, color = "blue") +
  geom_spatvector(data = vect(stic_locs_snapped), color = "black")+
  theme_void()

processed_folder <- "./7th_attempt/processed/"

writeRaster(outline_extended, paste0(processed_folder, "w3.tif"),overwrite=TRUE)
writeRaster(dem_final,paste0(processed_folder, "dem.tif"),overwrite=TRUE)
```

```{r claude channel fixer}

# Function to merge selected names into single row
merge_segments <- function(names_to_merge, sv_sf, snap_tolerance = 0.1){
  others   <- sv_sf[!sv_sf$ID %in% names_to_merge, ]
  to_merge <- sv_sf[ sv_sf$ID %in% names_to_merge, ]
  
  # Snap before union to close sub-mm endpoint gaps
  merged_geom <- st_snap(st_geometry(to_merge),
                          st_geometry(to_merge),
                          tolerance = snap_tolerance)
  
  # st_line_merge turns connected MULTILINESTRING into single LINESTRING
  merged_geom <- st_line_merge(st_union(merged_geom))
  
  if(st_geometry_type(merged_geom) == "MULTILINESTRING")
    warning(paste("Segments", paste(names_to_merge, collapse=","),
                  "did not form a single connected line — check they share endpoints",
                  "or increase snap_tolerance."))
  
  merged_row <- st_sf(ID = names_to_merge[1], geometry = merged_geom)
  rbind(merged_row, others)
}

# =============================================================================
# Helper: split a line at a fractional position along its length
# Returns a list of two sf lines (upstream part, downstream part)
# =============================================================================
split_line_at_fraction <- function(line_sf, frac){
  # frac in [0,1]: 0 = start, 1 = end of line
  # Uses lwgeom::st_linesubstring
  if(frac <= 0 || frac >= 1) return(list(line_sf, line_sf))
  
  part1 <- lwgeom::st_linesubstring(line_sf, 0,    frac)
  part2 <- lwgeom::st_linesubstring(line_sf, frac, 1)
  list(part1, part2)
}

# =============================================================================
# Add this helper inside build_dynatop_channel, after the splitting block
# and before "── 3. Build coordinate-based node topology ──"
# =============================================================================

merge_short_segments <- function(all_segs, min_length_m, node_tolerance){
  
  # Iteratively merge short segments into their nearest neighbour
  # until no segments below min_length_m remain
  max_iter <- 50L
  
  for(iter in seq_len(max_iter)){
    
    seg_lengths <- as.numeric(st_length(all_segs))
    short_idx   <- which(seg_lengths < min_length_m)
    
    if(length(short_idx) == 0){
      message(sprintf("  merge_short_segments: done after %d pass(es).", iter - 1))
      break
    }
    
    message(sprintf("  Pass %d: merging %d segment(s) < %g m.",
                    iter, length(short_idx), min_length_m))
    
    # Process shortest segment first
    target <- short_idx[which.min(seg_lengths[short_idx])]
    
    # Find topologically connected neighbours by shared endpoint
    precision  <- max(0, ceiling(-log10(node_tolerance)) + 1)
    
    get_endpoints <- function(geom){
      co <- st_coordinates(geom)
      list(
        start = round(co[1,        c("X","Y")], precision),
        end   = round(co[nrow(co), c("X","Y")], precision)
      )
    }
    
    ep_target <- get_endpoints(all_segs[target, ])
    
    # Find segments that share an endpoint with target (excluding itself)
    neighbours <- c()
    for(i in seq_len(nrow(all_segs))){
      if(i == target) next
      ep_i <- get_endpoints(all_segs[i, ])
      # Check all four endpoint combinations
      if(
        all(abs(ep_i$start - ep_target$start) < node_tolerance) ||
        all(abs(ep_i$start - ep_target$end)   < node_tolerance) ||
        all(abs(ep_i$end   - ep_target$start) < node_tolerance) ||
        all(abs(ep_i$end   - ep_target$end)   < node_tolerance)
      ){
        neighbours <- c(neighbours, i)
      }
    }
    
    if(length(neighbours) == 0){
      # Isolated segment — remove it (no valid neighbour to merge into)
      message(sprintf("  Removing isolated segment %d (no connected neighbour).", target))
      all_segs <- all_segs[-target, ]
      next
    }
    
    # Prefer merging into the shortest neighbour to keep segments balanced
    # (avoids creating one giant segment by always merging into the longest)
    best_neighbour <- neighbours[which.min(seg_lengths[neighbours])]
    
    # Merge geometries using st_line_merge(st_union(...)) for clean LINESTRING
    merged_geom <- st_line_merge(
      st_union(
        st_snap(
          c(st_geometry(all_segs[target, ]),
            st_geometry(all_segs[best_neighbour, ])),
          c(st_geometry(all_segs[target, ]),
            st_geometry(all_segs[best_neighbour, ])),
          tolerance = node_tolerance * 10
        )
      )
    )
    
    if(st_geometry_type(merged_geom) == "MULTILINESTRING"){
      # Snap tolerance wasn't enough — try larger snap and union
      merged_geom <- st_line_merge(
        st_union(st_geometry(all_segs[target, ]),
                 st_geometry(all_segs[best_neighbour, ]))
      )
    }
    
    # Replace neighbour row with merged geometry, remove target row
    all_segs[best_neighbour, ] <- st_sf(geometry = merged_geom)
    all_segs <- all_segs[-target, ]
    rownames(all_segs) <- NULL
  }
  
  if(iter == max_iter)
    warning("merge_short_segments hit max iterations — check for disconnected geometry.")
  
  all_segs
}

# =============================================================================
# FIXED build_dynatop_channel
# Key change: split points create a ±buffer_m segment unless bounded by
# an adjacent split or line endpoint
# =============================================================================

build_dynatop_channel <- function(
    channel_sv,
    split_points_sv,
    buffer_m       = 5,        # half-length of observation segments (m)
    min_length_m   = buffer_m * 0.8, # slivers below this are merged
    snap_tolerance = 1,        # max snap distance for points to line (m)
    node_tolerance = 0.01,     # coordinate precision for endpoint matching
    default_width  = 1,
    default_slope  = 0.001
){
  
  channel_sf <- st_cast(st_as_sf(channel_sv), "LINESTRING")
  points_sf  <- st_as_sf(split_points_sv)
  
  if(st_crs(channel_sf) != st_crs(points_sf)){
    message("CRS mismatch — reprojecting split points.")
    points_sf <- st_transform(points_sf, st_crs(channel_sf))
  }
  
  message(sprintf("Input: %d line feature(s), %d split point(s).",
                  nrow(channel_sf), nrow(points_sf)))
  
  # ── 1. Snap split points to nearest position on the channel ──────────────
  channel_union <- st_union(channel_sf)
  nearest_segs  <- st_nearest_points(points_sf, channel_union)
  snapped_pts   <- st_cast(nearest_segs, "POINT")[seq(2, 2*nrow(points_sf), 2)]
  
  snap_dists <- as.numeric(st_distance(
    points_sf,
    st_sf(geometry = snapped_pts),
    by_element = TRUE
  ))
  far <- which(snap_dists > snap_tolerance)
  if(length(far) > 0)
    warning(sprintf("%d point(s) >%g m from channel (rows: %s).",
                    length(far), snap_tolerance, paste(far, collapse=",")),
            call.=FALSE)
  
  # ── 2. For each line, find which points fall on it and where ─────────────
  # For each snapped point on a line, compute fractions for:
  #   frac_pt  — position of the point itself
  #   frac_lo  — frac_pt - buffer_m/line_length  (or 0 if too close to start)
  #   frac_hi  — frac_pt + buffer_m/line_length  (or 1 if too close to end)
  # All fractions for a line are sorted and de-duplicated before splitting.
  
  segment_list <- vector("list", nrow(channel_sf))
  
  for(i in seq_len(nrow(channel_sf))){
    this_line    <- channel_sf[i, ]
    line_length  <- as.numeric(st_length(this_line))
    
    dist_to_line <- as.numeric(st_distance(
      st_sf(geometry = snapped_pts), this_line
    ))
    on_this <- which(dist_to_line <= node_tolerance * 10)
    
    if(length(on_this) == 0){
      # No split points on this line — keep as-is
      segment_list[[i]] <- st_sf(geometry = st_geometry(this_line))
      next
    }
    
    # Compute fractional positions of each point along this line
    fracs_all <- c(0, 1)  # always include endpoints
    
    for(j in on_this){
      pt <- snapped_pts[j]
      
      # Project point onto line to get distance from start
      # Use st_nearest_points to find the foot of perpendicular
      nearest <- st_nearest_points(pt, st_geometry(this_line))
      foot    <- st_cast(nearest, "POINT")[2]
      
      # Distance from line start to foot
      # Compute by splitting line at foot and measuring first part
      # Use parameterized position via lwgeom
      frac_pt <- tryCatch({
        # Line substring from 0 to foot: measure its length / total length
        d_from_start <- as.numeric(
          st_length(lwgeom::st_linesubstring(this_line, 0,
            # Binary search for fraction — lwgeom doesn't give inverse directly
            # Use st_line_sample approximation then refine
            {
              # Approximate: sample many points, find closest
              n_samp  <- max(100, ceiling(line_length))
              fracs_s <- seq(0, 1, length.out = n_samp)
              samples <- st_line_sample(this_line, sample = fracs_s)
              samples_pts <- st_cast(samples, "POINT")
              dists   <- as.numeric(st_distance(foot, samples_pts))
              fracs_s[which.min(dists)]
            }
          ))
        )
        d_from_start / line_length
      }, error = function(e) NA_real_)
      
      if(is.na(frac_pt) || frac_pt < 0 || frac_pt > 1) next
      
      # ±buffer_m in fractional units, clamped to [0,1]
      frac_lo <- max(0, frac_pt - buffer_m / line_length)
      frac_hi <- min(1, frac_pt + buffer_m / line_length)
      
      # Only add interior fractions (not already at 0 or 1)
      if(frac_lo > node_tolerance)             fracs_all <- c(fracs_all, frac_lo)
      if(frac_hi < 1 - node_tolerance)         fracs_all <- c(fracs_all, frac_hi)
    }
    
    # Sort, deduplicate close fractions, split
    fracs_all <- sort(unique(fracs_all))
    fracs_all <- fracs_all[c(TRUE, diff(fracs_all) > node_tolerance / line_length)]
    
    # Build sub-segments from consecutive fraction pairs
    parts <- vector("list", length(fracs_all) - 1)
    for(k in seq_len(length(fracs_all) - 1)){
      f1 <- fracs_all[k]
      f2 <- fracs_all[k + 1]
      parts[[k]] <- tryCatch(
        lwgeom::st_linesubstring(this_line, f1, f2),
        error = function(e){
          warning(sprintf("st_linesubstring failed on line %d [%.4f, %.4f]", i, f1, f2))
          NULL
        }
      )
    }
    
    parts <- Filter(Negate(is.null), parts)
    if(length(parts) == 0){
      segment_list[[i]] <- st_sf(geometry = st_geometry(this_line))
    } else {
      geoms <- do.call(c, lapply(parts, st_geometry))
      segment_list[[i]] <- st_sf(geometry = geoms)
    }
  }
  
  all_segs <- do.call(rbind, segment_list)
  rownames(all_segs) <- NULL
  
  # Remove degenerate zero-length segments
  seg_lengths <- as.numeric(st_length(all_segs))
  all_segs    <- all_segs[seg_lengths > node_tolerance, ]
  n_segs      <- nrow(all_segs)
  
  message(sprintf("After splitting at ±%gm buffers: %d segment(s).", buffer_m, n_segs))
  
  
  
  # ── 2b. Remove short remnant segments ────────────────────────────────────
  # min_length_m: segments shorter than this are merged into a neighbour.
  # Set to slightly less than buffer_m so that genuine sensor segments
  # (≈ 2*buffer_m = 10m) are never merged away, but slivers are.
  min_length_m <- buffer_m * 0.8   # e.g. 4m for buffer_m=5

  n_before <- nrow(all_segs)
  all_segs <- merge_short_segments(all_segs, min_length_m, node_tolerance)
  n_after  <- nrow(all_segs)
  
  message(sprintf("Short segment removal: %d → %d segment(s) (removed %d).",
                  n_before, n_after, n_before - n_after))
  
  
  # ── 3. Build coordinate-based node topology ───────────────────────────────
  precision <- max(0, ceiling(-log10(node_tolerance)) + 1)
  
  get_endpoints <- function(geom){
    co <- st_coordinates(geom)
    rbind(
      round(co[1,        c("X","Y")], precision),
      round(co[nrow(co), c("X","Y")], precision)
    )
  }
  
  ep_matrix   <- do.call(rbind, lapply(seq_len(n_segs),
                                        function(i) get_endpoints(all_segs[i,])))
  coord_keys  <- paste(ep_matrix[,"X"], ep_matrix[,"Y"])
  unique_keys <- unique(coord_keys)
  node_ids    <- match(coord_keys, unique_keys)
  start_ids   <- node_ids[seq(1, 2*n_segs, 2)]
  end_ids     <- node_ids[seq(2, 2*n_segs, 2)]
  
  # ── 4. Topological sort — BFS from outlet ─────────────────────────────────
  sn_to_row <- setNames(seq_len(n_segs), as.character(start_ids))
  ds        <- unname(sn_to_row[as.character(end_ids)])  # NA = outlet
  
  outlets <- which(is.na(ds))
  if(length(outlets) == 0)
    stop("No outlet found. Check that lines are directed upstream→downstream.")
  if(length(outlets) > 1)
    warning(sprintf("%d outlets detected — network may be disconnected.",
                    length(outlets)), call.=FALSE)
  
  upstream_of <- vector("list", n_segs)
  for(i in seq_len(n_segs))
    if(!is.na(ds[i])) upstream_of[[ds[i]]] <- c(upstream_of[[ds[i]]], i)
  
  visited   <- logical(n_segs)
  new_order <- integer(0)
  queue     <- outlets
  while(length(queue) > 0){
    cur <- queue[1]; queue <- queue[-1]
    if(!visited[cur]){
      visited[cur] <- TRUE
      new_order    <- c(new_order, cur)
      queue        <- c(queue, upstream_of[[cur]])
    }
  }
  
  unreachable <- setdiff(seq_len(n_segs), new_order)
  if(length(unreachable) > 0)
    warning(sprintf("%d segment(s) unreachable from outlet — check network connectivity.",
                    length(unreachable)), call.=FALSE)
  
  n_sorted <- length(new_order)
  
  # ── 5. Assign topological ids ─────────────────────────────────────────────
  topo_id              <- seq_len(n_sorted)
  sorted_start_orig    <- start_ids[new_order]
  sorted_end_orig      <- end_ids[new_order]
  sn_orig_to_sorted    <- setNames(seq_len(n_sorted),
                                    as.character(sorted_start_orig))
  ds_sorted            <- unname(sn_orig_to_sorted[as.character(sorted_end_orig)])
  
  new_startNode <- topo_id
  new_endNode   <- ifelse(is.na(ds_sorted), 0L, topo_id[ds_sorted])
  
  violations <- which(new_endNode != 0 & new_endNode >= new_startNode)
  if(length(violations) > 0)
    stop(sprintf("Topological id violations at sorted rows: %s.",
                 paste(violations, collapse=", ")))
  
  sorted_lengths <- as.numeric(st_length(all_segs[new_order, ]))
  
  # ── 6. Assemble output ────────────────────────────────────────────────────
  channel_df <- data.frame(
    name      = topo_id,
    startNode = new_startNode,
    endNode   = new_endNode,
    length    = sorted_lengths,
    width     = default_width,
    slope     = default_slope,
    area      = sorted_lengths * default_width,
    stringsAsFactors = FALSE
  )
  
  sorted_geom    <- st_geometry(all_segs[new_order, ])
  channel_sf_out <- st_sf(channel_df, geometry = sorted_geom)
  channel_sv_out <- vect(channel_sf_out)
  crs(channel_sv_out) <- crs(channel_sv)
  
  validate_channel_sv(channel_sv_out, verbose = TRUE)
  
  # ── 7. Run convert_channel ────────────────────────────────────────────────
  conv_channel <- convert_channel(
    channel_sv_out,
    property_names = c(
      name      = "name",
      length    = "length",
      startNode = "startNode",
      endNode   = "endNode"
    ),
    default_width = default_width,
    default_slope = default_slope
  )
  
  list(
    channel_sv   = channel_sv_out,
    channel_df   = channel_df,
    conv_channel = conv_channel
  )
}

# =============================================================================
# Validation helper (unchanged)
# =============================================================================

validate_channel_sv <- function(sv, verbose = TRUE){
  df          <- as.data.frame(sv)
  outlet_rows <- df$endNode == 0 | !df$endNode %in% df$startNode
  bad         <- which(!outlet_rows & df$endNode >= df$startNode)
  if(verbose){
    message(sprintf(
      "Segments: %d | Outlet startNode(s): %s | Violations: %d",
      nrow(df),
      paste(df$startNode[outlet_rows], collapse=", "),
      length(bad)
    ))
    if(length(bad) > 0){
      message("Violations:")
      print(df[bad, c("name","startNode","endNode")], row.names=FALSE)
    } else {
      message("PASS — ordering correct for convert_channel and dynatop.")
    }
  }
  invisible(length(bad) == 0)
}

# =============================================================================
# USAGE — adapt paths and IDs to your watershed
# =============================================================================

# ── Step 1: Remove non-existent channels ─────────────────────────────────────

locs <- vect(stic_locs_snapped)

## They've been trying to stop the wave, but the wave don't stop
channel <- vect(w3_streams_vect, crs = crs1) %>% 
  crop(w3_outline) %>% 
    mutate(ID = seq(1, n(), 1))


ggplot()+
  geom_spatvector(data = channel)+
  geom_spatvector_text(data = channel, aes(label = ID))+
  geom_spatvector(data = locs, color = "blue")

channel_trimmed <- channel %>%
  filter(!ID %in% c(1,2,8,9,4,14,30,35,45,24,27,16,12))
# 10 m resolution trimmed: c(1,5,6,4,12,7,23,35,32,41)
# 3 m resolution trimmed: c(1, 8,9, 12, 19, 4, 39, 25, 15, 2)

ggplot()+
  geom_spatvector(data = channel_trimmed)+
  geom_spatvector_text(data = channel_trimmed, aes(label = ID))+
    geom_spatvector(data = locs, color = "blue")
# ── Step 2: Convert to sf ────────────────────────────────────────────────────
sv_sf <- st_as_sf(channel_trimmed %>% select(ID))

# ── Step 3: Merge segments that should be one reach ──────────────────────────
# IMPORTANT: merge BEFORE renumbering — IDs must exist in sv_sf
# Use increased snap_tolerance if segments don't share exact endpoints

test <- merge_segments(names_to_merge = c(3,7,6),
                       sv_sf = sv_sf,
                       snap_tolerance = 0.1)
ggplot()+
  geom_spatvector(data = test)+
  geom_spatvector_text(data = test, aes(label = ID))+
    geom_spatvector(data = locs, color = "blue")

vect_final <- vect(test) %>% 
  # renumber segments
  mutate(ID = row_number())

ggplot()+
  geom_spatvector(data = vect_final)+
  geom_spatvector_text(data = vect_final, aes(label = ID))+
    geom_spatvector(data = locs, color = "blue")

# ── Step 6: Build channel with observation point splits ──────────────────────
# split_points_sv  — your STIC/observation locations (already projected)
# buffer_m = 5     — creates a 10m segment (5m each side) at each location
#                    unless bounded by an adjacent split or line endpoint
result <- build_dynatop_channel(
  channel_sv      = vect_final,
  split_points_sv = locs,
  buffer_m        = 5,          # ← 5m either side of observation point
  min_length_m   = buffer_m * 2,
  snap_tolerance  = 20,         # max distance to snap obs point to channel
  node_tolerance  = 0.01,
  default_width   = 1,
  default_slope   = 0.001
)

# ── Step 7: Inspect ───────────────────────────────────────────────────────────
print(result$channel_df)

ggplot() +
  geom_sf(data = st_as_sf(result$channel_sv)) +
    geom_sf(data = locs, color = "blue", size = 2) +
  geom_sf_text(data = st_as_sf(result$channel_sv), aes(label = name), size = 3) +
  ggtitle("final channel HRUs")+
  theme_void()

validate_channel_sv(result$channel_sv)

writeVector(
  result$conv_channel,
  paste0(processed_folder, "channel.shp"),
  overwrite = TRUE
)


```


```{r identify ID to hru ID}
# =============================================================================
# Extract sensor-to-HRU mapping from build_dynatop_channel result
# =============================================================================
get_sensor_hru_map <- function(channel_sv, split_points_sv, 
                                buffer_m = 5,
                                node_tolerance = 0.01){
  
  channel_sf <- st_as_sf(channel_sv)
  points_sf  <- st_as_sf(split_points_sv)
  
  if(st_crs(channel_sf) != st_crs(points_sf))
    points_sf <- st_transform(points_sf, st_crs(channel_sf))
  
  # Snap points to channel (same logic as build_dynatop_channel)
  channel_union <- st_union(channel_sf)
  nearest_segs  <- st_nearest_points(points_sf, channel_union)
  snapped_pts   <- st_cast(nearest_segs, "POINT")[seq(2, 2*nrow(points_sf), 2)]
  
  # For each snapped point, find which channel segment it falls within
  # A sensor's segment is the one whose midpoint is closest to the snapped point
  # AND whose length is <= 2*buffer_m + small tolerance (i.e. it's a split segment)
  seg_lengths   <- as.numeric(st_length(channel_sf))
  seg_midpoints <- st_centroid(channel_sf)
  
  mapping <- do.call(rbind, lapply(seq_len(nrow(points_sf)), function(i){
    
    pt <- snapped_pts[i]
    
    # Distance from snapped point to each segment midpoint
    dists <- as.numeric(st_distance(pt, seg_midpoints))
    
    # Candidate segments: those short enough to be a sensor segment
    # (length <= 2*buffer_m + tolerance) AND closest to the snapped point
    is_sensor_seg <- seg_lengths <= (2 * buffer_m + 1)  # 1m tolerance
    
    if(any(is_sensor_seg)){
      # Among sensor-length segments, pick closest midpoint
      candidate_dists <- ifelse(is_sensor_seg, dists, Inf)
      best_idx <- which.min(candidate_dists)
    } else {
      # Fallback: just pick closest segment overall
      best_idx <- which.min(dists)
    }
    
    # Get sensor ID from the split_points_sv attribute
    # Tries common ID column names — adapt if yours differs
    id_cols <- c("number", "ID", "id", "sensor_id", "name", "sensor_name")
    id_col  <- id_cols[id_cols %in% names(points_sf)][1]
    sensor_id <- if(!is.na(id_col)) points_sf[[id_col]][i] else i
    
    data.frame(
      sensor_id  = sensor_id,
      hru_id     = channel_sf$startNode[best_idx],  # startNode IS the HRU id
      hru_name   = channel_sf$name[best_idx],
      seg_length = round(seg_lengths[best_idx], 2),
      snap_dist  = round(as.numeric(st_distance(
                     points_sf[i,], st_sf(geometry = snapped_pts[i])
                   )), 2),
      stringsAsFactors = FALSE
    )
  }))
  
  rownames(mapping) <- NULL
  return(mapping)
}

# ── Usage ─────────────────────────────────────────────────────────────────────
locs <- vect(stic_locs_snapped)

vect_final <- vect("./7th_attempt/processed/channel.shp")
plot(vect_final)

ggplot() +
  geom_sf(data = st_as_sf(vect_final)) +
    geom_sf(data = locs, color = "blue", size = 2) +
  geom_sf_text(data = st_as_sf(vect_final), aes(label = name), size = 3) +
  ggtitle("final channel HRUs")+
  theme_void()

sensor_hru_map <- get_sensor_hru_map(
  channel_sv          = vect_final,
  split_points_sv = locs,
  buffer_m        = 5
)


print(sensor_hru_map)
sensor_hru_map$sensor_id <- as.numeric(sensor_hru_map$sensor_id)

```



## TIMESERIES
```{r}
start <- "2020-01-01"
end <- "2024-11-26"
```

```{r discharge}


# discharge every 15 mins
flow_15min <- read_csv("https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/1/18/1b101627265d0b5763fa0bba93d93bf7") 
q_24 <- read_csv("/Users/johnmorgan/Documents/VT Research/HBTopModel/w3_discharge_24.csv") %>% 
  mutate(datetime = mdy_hm(datetime)) %>% 
  filter(datetime > max(flow_15min$DATETIME)) %>% 
  mutate(WS = 3,
         Flag = "P",
         Discharge_ls = 28.3168 * Q_cfs) %>% 
  rename(DATETIME = datetime) %>% 
  select(WS, DATETIME, Discharge_ls, Flag)
  

flow_xts <- 
  flow_15min %>%
  rbind(q_24) %>% 
  filter(DATETIME >= start,
         DATETIME <= end) %>% 
  # Create a new column that groups time into 15-minute buckets
  mutate(time_15 = floor_date(DATETIME, "15 mins")) %>%
  # Group by those buckets
  group_by(time_15) %>%
  # Average the 5-minute instantaneous rates to get the 15-minute rate
  summarize(Discharge_ls = mean(Discharge_ls, na.rm = TRUE)) %>%
  # convert to m3_s, or should it be m3/day?>
  mutate(Discharge_m3_s = Discharge_ls / 1000) %>% 
  as.xts()

flow_xts <- 
  flow_15min %>%
  rbind(q_24) %>% 
  filter(DATETIME >= start,
         DATETIME <= end) %>% 
  # Create a new column that groups time into 15-minute buckets
  mutate(time_hr = floor_date(DATETIME, "1 hour")) %>%
  # Group by those buckets
  group_by(time_hr) %>%
  # Average the 5-minute instantaneous rates to get the 15-minute rate
  summarize(Discharge_ls = mean(Discharge_ls, na.rm = TRUE)) %>%
  # convert to m3_s, or should it be m3/day?>
  mutate(Discharge_m3_s = Discharge_ls / 1000) %>% 
  as.xts()

```
Units of precipitation and pet are expected to be m accrued over the previous timestep.
```{r precipitation}
precip <- read_csv("https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/278/7/132ffe0622ff3f30505c774d24a6b65d")
# precip in mm accrued over timestep, needs to be m accrued over time step

precip_xts <- 
  precip %>% 
  filter(DateTime >= start,
         DateTime <= end) %>% 
  select(DateTime, precip) %>% 
  mutate(precip = precip / 1000) %>% # convert to m
  # Create a new column that groups time into 15-minute buckets
  mutate(time_hr = floor_date(DateTime, "1 hour")) %>%
  # Group by those buckets
  group_by(time_hr) %>%
  # Average the 5-minute instantaneous rates to get the 15-minute rate
  summarize(precip = sum(precip, na.rm = TRUE)) %>%
  as.xts()
```
```{r pet}
# air temp already 15 min increment
air <- read_csv("https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/322/7/a96172a26f381b9ec7005144d12e0820")

air_formatted <- air %>% 
  filter(DateTime >= start,
         DateTime <= end) %>% 
  select(DateTime, airTemp) 

# convert airTemp to pet with hargreaves

lat <- 45   # latitude in decimal degrees (e.g., Blacksburg VA)

# --- Hargreaves PET -----------------------------------------------------

hargreaves_pet <- function(date, temp, lat){

  J <- yday(date)                 # day of year
  phi <- lat * pi/180             # latitude radians
  
  # extraterrestrial radiation (Ra)
  dr  <- 1 + 0.033*cos(2*pi*J/365)
  delta <- 0.409*sin(2*pi*J/365 - 1.39)
  ws <- acos(-tan(phi)*tan(delta))

  Gsc <- 0.0820   # MJ m-2 min-1
  
  Ra <- (24*60/pi) * Gsc * dr *
        (ws*sin(phi)*sin(delta) +
         cos(phi)*cos(delta)*sin(ws))

  PET <- 0.0023 * Ra *
         (temp + 17.8) *
         sqrt(pmax(temp, 0))

  return(PET)   # mm/day
}

# --- Calculate PET ------------------------------------------------------

air_formatted$PET_mm_day <- hargreaves_pet(
  air_formatted$DateTime,
  air_formatted$airTemp,
  lat
)


pet_xts <- 
  air_formatted %>% 
  select(DateTime, PET_mm_day, airTemp) %>% 
  mutate(PET_m_day = PET_mm_day / (1000 * 96)) %>% # convert to m, and 15 min
  rename(pet = PET_m_day) %>% 
  # Create a new column that groups time into 15-minute buckets
  mutate(time_hr = floor_date(DateTime, "1 hour")) %>%
  # Group by those buckets
  group_by(time_hr) %>%
  # Average the 5-minute instantaneous rates to get the 15-minute rate
  summarize(pet = mean(pet, na.rm = TRUE),
            airTemp = mean(airTemp, na.rm = TRUE)) %>%
  #select() %>% 
  as.xts()

 
```
```{r combine}

obs <- merge(flow_xts[,2], # flow does not go beyond 2024-05-31 23:45:00
             precip_xts,
             pet_xts,
             all=FALSE) ## merge by time stamp

names(obs) <- c("Discharge_m3_s", "precip","pet","temp_series" )
```

## CREATE DYNATOP OBJECT
```{r create object}
#devtools::install_github("waternumbers/dynatopGIS")
#library(dynatopGIS)

ctch <- dynatopGIS$new(file.path(".","7th_attempt", "dynaGISgoodChannel7"))
# add catchment, DEM, channel
ctch$add_catchment(paste0(processed_folder, "w3.tif"))
ctch$add_dem(paste0(processed_folder, "dem.tif"))
ctch$add_channel( paste0(processed_folder, "channel.shp"))

ctch$sink_fill()

# bands
ctch$compute_band()
ctch$plot_layer("band")
# twi
ctch$compute_properties()
ctch$plot_layer('atb')
# flow lengths
ctch$compute_flow_lengths(flow_routing="shortest")
ctch$plot_layer("shortest_flow_length")

```

```{r bring in other datasets}

# soil map
soil <- terra::rast("https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/380/2/f083871fa0116b17b7398ca48eac7627")
soil_factor <- as.factor(soil)

soil_r  <- resample(project(soil_factor, dem_crop_mask), dem_crop_mask, method="mode")

crop_soil <- crop(soil_r, dem_crop_mask)
mask_soil <- mask(soil_r, dem_crop_mask)
plot(mask_soil)


key <- read_csv("/Users/johnmorgan/Documents/soil_depth_planning/HPU_num_key.csv")


levels(mask_soil) <- data.frame(
  "ID" = c(
    5, 4, 2, 8, 3, 1, 6, 7
  ),
  "f083871fa0116b17b7398ca48eac7627" = c(
  "5"," 4"," 2", "8", "3", "1", "6", "7")
)

colors <- c("black",
            "#CCCCCC",
            "#734C00",
            "#FFFF73",
            "#267300",
            "#55FF00",
            "black"
            )
soil_names <- c(
  "Bedrock Histosol",
  "E",
  "Bhs",
  "Typical",
  "Bimodal",
  "Bh"
)
ggplot() +
    geom_spatraster(data = (mask_soil))+
scale_fill_manual(values = colors,
                  labels = soil_names,
                  na.value = NA)+
  theme_void()+
      geom_sf(data = w3_outline, colour = "black", lwd = 1, fill = NA) +
  labs(title = "1 m overburden thickness contours")



# read in krig of depth to bedrock
krig_location <- "/Users/johnmorgan/Documents/soil_depth_planning/krig_output.tif"
k_terra <- rast(krig_location)
plot(k_terra)

# resample to the same pixel size and alignment
depth  <- resample(project(k_terra, dem_crop_mask), dem_crop_mask, method="bilinear") %>% 
  crop(dem_crop_mask) %>% 
  mask(dem_crop_mask)

# check all of their extents
plot(dem_crop_mask)
plot(depth)
plot(mask_soil)


# extend both rasters by 1 pixel
depth_ext <- extend(depth,1)
soil_ext <- extend(mask_soil,1)

# check all of their extents
plot(depth_ext)
plot(soil_ext)
#ctch$plot_layer("dem", add_channel=TRUE)

reclass_matrix <- matrix(c(1, 1,
                           2, 3,
                           3, 2,
                           4, 3,
                           5, 3,
                           7, 3,
                           8, 2), 
                         ncol = 2, byrow = TRUE)

# Apply classification
soil_ext <- classify(soil_ext, reclass_matrix)
plot(soil_ext)


#ctch$add_layer(depth_ext, layer_name = "depth")
ctch$add_layer(soil_ext, layer_name = "soil")
#ctch$classify("depth_5","depth",5)
#ctch$plot_layer("depth_5")

```


```{r classifying and creating model}
## simple equally spaced cuts of TWI
ctch$classify("atb_20","atb",20)
ctch$plot_layer("atb_20")

ctch$combine_classes("atb20_soil",c("atb_20","soil"))
ctch$plot_layer("atb20_soil")


ctch$create_model(file.path(".", "7th_attempt", "dynaGISgoodChannel7","model_preSnow"), # name of new model
                  "atb20_soil", # classification to base model on
                  sf_opt="cnst",
                  sz_opt="exp"
                  )

# read in model to modify for snow routine
mdl <- readRDS( file.path(".", "7th_attempt", "dynaGISgoodChannel7","model_preSnow.rds")) ## read the model in

```

```{r fix hru topology}
fix_hru_topology <- function(mdl, verbose = TRUE){
  hru <- mdl$hru
  
  # ── Rename helper — updates id AND all references ────────────────────────
  rename_hru <- function(hru, old_id, new_id){
    for(ii in seq_along(hru)){
      if(hru[[ii]]$id == old_id) hru[[ii]]$id <- new_id
      hru[[ii]]$sf_flow_direction$id[
        hru[[ii]]$sf_flow_direction$id == old_id] <- new_id
      hru[[ii]]$sz_flow_direction$id[
        hru[[ii]]$sz_flow_direction$id == old_id] <- new_id
    }
    return(hru)
  }
  
  # ── Find violations ───────────────────────────────────────────────────────
  find_violations <- function(hru){
    out <- do.call(rbind, lapply(seq_along(hru), function(ii){
      h      <- hru[[ii]]
      bad_sf <- h$sf_flow_direction$id[h$sf_flow_direction$id >= h$id]
      bad_sz <- h$sz_flow_direction$id[h$sz_flow_direction$id >= h$id]
      if(length(bad_sf) > 0 || length(bad_sz) > 0)
        data.frame(hru_id  = h$id,
                   max_bad = max(c(bad_sf, bad_sz)),
                   stringsAsFactors = FALSE)
    }))
    out
  }
  
  # ── Iterative fix ─────────────────────────────────────────────────────────
  # Strategy: find HRU X that flows to Y where Y >= X.
  # X must have id > Y (and > everything else X flows to).
  # New id for X = max(all downstream ids of X) + 1.
  # Every HRU currently occupying ids between old_X and new_X shifts down by 1
  # to keep the sequence contiguous.
  
  tmp_id  <- 9999L  # safe scratch id above any real id
  max_iter <- 50L
  
  for(iter in seq_len(max_iter)){
    violations <- find_violations(hru)
    if(is.null(violations) || nrow(violations) == 0){
      if(verbose) cat(sprintf("All violations resolved after %d iteration(s).\n",
                              iter - 1))
      break
    }
    if(verbose){
      cat(sprintf("Iteration %d — violations:\n", iter))
      print(violations)
    }
    
    # Fix the first violation
    v      <- violations[1, ]
    hru_id <- v$hru_id
    
    # Collect ALL downstream ids for this HRU (sf and sz)
    ii      <- which(sapply(hru, function(h) h$id == hru_id))
    all_dst <- c(hru[[ii]]$sf_flow_direction$id,
                 hru[[ii]]$sz_flow_direction$id)
    all_dst <- all_dst[!is.na(all_dst) & length(all_dst) > 0]
    
    # New id must be strictly above all downstream ids
    new_id  <- max(all_dst) + 1L
    old_id  <- hru_id
    
    if(verbose)
      cat(sprintf("  Moving HRU %d -> %d (flows to: %s)\n",
                  old_id, new_id, paste(all_dst, collapse=",")))
    
    # Step 1: park the offending HRU at tmp_id
    hru <- rename_hru(hru, old_id, tmp_id)
    
    # Step 2: shift every HRU with id in (old_id, new_id] down by 1
    # Do this from lowest to highest to avoid collisions
    ids_to_shift <- sort(sapply(hru, function(h) h$id))
    ids_to_shift <- ids_to_shift[ids_to_shift > old_id &
                                   ids_to_shift <= new_id &
                                   ids_to_shift != tmp_id]
    
    for(shift_id in ids_to_shift){
      hru <- rename_hru(hru, shift_id, shift_id - 1L)
    }
    
    # Step 3: move parked HRU from tmp_id to new_id
    hru <- rename_hru(hru, tmp_id, new_id)
  }
  
  if(iter == max_iter){
    remaining <- find_violations(hru)
    if(!is.null(remaining) && nrow(remaining) > 0){
      warning("Could not resolve all violations after ", max_iter,
              " iterations. Remaining:\n")
      print(remaining)
      cat("This may indicate a cycle in the channel network.\n")
    }
  }
  
  # ── Verify contiguous sequence ────────────────────────────────────────────
  final_ids <- sort(sapply(hru, function(h) h$id))
  expected  <- seq(min(final_ids), max(final_ids))
  gaps      <- setdiff(expected, final_ids)
  if(length(gaps) > 0){
    warning("Gaps in id sequence after fix: ", paste(gaps, collapse=", "))
  } else {
    if(verbose) cat("ID sequence contiguous: OK\n")
  }
  
  mdl$hru <- hru
  return(mdl)
}

mdl <- fix_hru_topology(mdl = mdl, verbose = TRUE)
```

```{r identify sensor locations amongst channel HRUs}
# actually need to figure out which HRU in final model corresponds with sensor location

# write function to filter HRU object for HRU's with a specific class name.
sensor_hru_map
mdl$hru[[1]]$class$name

hid <- sapply(mdl$hru,function(h){h$id})
## create a vector of the channel names with as storaged in the class variabel
cname <- sapply(mdl$hru,function(h){h$class$name})
idx <- match(sensor_hru_map$hru_name,cname) ## index of HRU's corresponding to gauges
## create as a data.frame
sensor_hru_key <- data.frame(sensor_id = sensor_hru_map$sensor_id,
           hru_id = as.integer(hid[idx]),
           name = sensor_hru_map$hru_name
                           )
```

```{r identify soil groups}
# =============================================================================
# EXTRACT SOIL GROUPS FOR HILLSLOPE HRUs
# =============================================================================
# Assuming 'soil' was used to generate your HRU classes and is stored in $class
# You may need to change `h$class$soil` to wherever your class name is stored 
# (e.g., h$properties["soil"] or similar).

hru_soil_map <- lapply(seq_along(mdl$hru), function(ii) {
  h <- mdl$hru[[ii]]
  is_channel <- !is.na(h$class$endNode)
  
  if (!is_channel) {
    # CHANGE THIS to the exact location of your soil identifier
    soil_id <- as.character(h$class$soil) 
    data.frame(hru_index = ii, hru_id = h$id, soil_group = soil_id)
  } else {
    NULL
  }
})

hru_soil_map <- do.call(rbind, hru_soil_map)

# Map the raw soil names/IDs to a simple index (1, 2, 3) for parameter matching
unique_soils <- unique(hru_soil_map$soil_group)
hru_soil_map$soil_idx <- match(hru_soil_map$soil_group, unique_soils)

cat("Found", length(unique_soils), "soil groups across", nrow(hru_soil_map), "hillslope HRUs.\n")
```

```{r add snow parameters to model object}
# 2. Patch the HRUs with snow parameters, temp linkages, and expanded states
for (i in seq_along(mdl$hru)) {
  
  # A. Add the snow parameters
  mdl$hru[[i]]$snow <- list(
    type = "hbv",
    parameters = c(T_T = 0.0, C_f = 2.5, C_wh = 0.1)
  )
  
  # B. Add the temp linkage 
  # Note: The original precip linkage uses 'name' and 'fraction'
  # We will map it to a dummy "temp_series"
  mdl$hru[[i]]$temp <- list(
    name = "temp_series",
    fraction = 1.0
  )
  
  # C. Expand the states to include s_snow and s_liquid
  mdl$hru[[i]]$states <- c(
    mdl$hru[[i]]$states, 
    s_snow = 0.0, 
    s_liquid = 0.0
  )
}

# ensure that model has snow parameters included
```

## SAVE EVERYTHING IN ONE PLACE
```{r}
# indexes and keys
write_csv(sensor_hru_key, "./7th_attempt/sensor_hru_key.csv")
write_csv(hru_soil_map, "./7th_attempt/hru_soil_map.csv")

# channel network saved already


# observations
saveRDS(obs, file.path(".","7th_attempt", "obs.rds"))

# model
saveRDS(mdl, file.path(".", "7th_attempt", "dynaGISgoodChannel7","model_ready.rds"))

```


# ACTUALLY RUN MODEL
## READ IN STUFF
```{r}
## load the model
rm(mdl)
rm(obs)
mdl <- readRDS( file.path(".", "7th_attempt", "dynaGISgoodChannel7","model_ready.rds")) ## read the model in
obs <- readRDS(file.path(".", "7th_attempt", "obs.rds"))
vect_final <- vect("./7th_attempt/processed/channel.shp")
```

## TEST RUN
```{r test run model}

# set initial conditions
r0 <- as.numeric(obs$Discharge_m3_s[1])/ sum( sapply( mdl$hru, function(x){x$properties["area"]} ) )
mdl$hru <- lapply(mdl$hru,function(h){h$initialisation["r_uz_sz_0"] <- r0;h})


Sys.time()
ctch_mdl <- dynatop$new(mdl$hru, map = mdl$map
  )$add_data( obs
  )$initialise(
  )$sim(mdl$output_flux, sub_step = 3600
  )


mdl_after <- ctch_mdl$get_model()

mdl_after[[1]]$states
# Gauge Flows
sim_gauge <- ctch_mdl$get_output() ## extract the flow as an xts object
sim_gauge[1:50,]
plot(sim_gauge[,1],main="Simulated and observed flows",
     legend.loc='topright',
     xlim = c( as.POSIXct("2020-01-03", tz = "UTC")))
plot(merge(sim_gauge[,1], obs$Discharge_m3_s),main="Simulated and observed flows",
     legend.loc='topright')

kge(sim_gauge[,1], obs$Discharge_m3_s)
```

## SIMPLE PARAMETERIZE


```{r create pa_obs}
# =============================================================================
# 2. FLOW PRESENCE/ABSENCE OBSERVATIONS
# =============================================================================
#all_24 <- read_csv("/Users/johnmorgan/Documents/VT Research/HBTopModel/STIC_Synthesis/02_BinaryMatrix/Binary_hb_w3.csv")
all_24 <- read_csv("/Users/johnmorgan/Documents/VT Research/HBTopModel/STIC_Synthesis/all_24.csv") %>% 
  filter(wshed == "W3") %>% 
  select(datetime, wetdry, number, lat, long) %>% 
  filter(number %in% sensor_hru_key$sensor_id)



all_23 <- read_csv("/Users/johnmorgan/Documents/VT Research/HBTopModel/STIC_Synthesis/all_23.csv") %>% 
  filter(wshed == "W3") %>% 
  select(datetime, wetdry, ID, lat, long) %>% 
  rename(number = ID) %>% 
  filter(number %in% sensor_hru_key$sensor_id)



stic_locs <- all_23 %>% 
  bind_rows(all_24) %>%
  select(number, lat, long) %>%
  distinct() %>%
  vect(geom = c("long", "lat"), crs = "EPSG:4326") %>%
  project("EPSG:26919")

all_out <- do.call(rbind, lapply(seq_len(nrow(stic_locs)), function(i){
  dist_table <- distance(stic_locs[i,], vect_final)
  idx        <- which.min(dist_table)[1]
  data.frame(ID         = stic_locs[i,]$number,
             channel_ID = vect_final[idx,]$name,
             dist       = min(dist_table))
}))

sensor_hru_key
pa_obs <- 
  all_23 %>% 
  bind_rows(all_24) %>%
  filter(number %in% sensor_hru_key$sensor_id) %>%
  select(datetime, number, wetdry) %>%
  rename(ID = number) %>%
  left_join(sensor_hru_key, by = c("ID" = "sensor_id")) %>%
  #left_join(chosen, by = "ID", relationship = "many-to-one") %>%
  mutate(
    observed = as.integer(case_when(wetdry == "wet" ~ 1, wetdry == "dry" ~ 0)),
    datetime = as.POSIXct(datetime, tz = "UTC")
  ) %>%
  select(hru_id, datetime, observed) %>%
  filter(datetime >= cal_start, datetime <= cal_end,
         !is.na(hru_id), !is.na(observed))

cat("PA sites:", sort(unique(pa_obs$hru_id)), "\n")
cat("PA obs:", nrow(pa_obs), "| flowing:", sum(pa_obs$observed == 1),
    "| dry:", sum(pa_obs$observed == 0), "\n")

# save as csv
write_csv(pa_obs, "./7th_attempt/pa_obs.csv")
```



# PARAMETERIZE

```{r param attempt 3}
# =============================================================================
# 1. LOAD MODEL AND DATA
# =============================================================================
# mdl <- readRDS(file.path(".", "hb_data", "dynaGIS", "atb_20.rds"))
# obs <- readRDS(file.path(".", "processed", "obs.rds"))

# -- Windows ------------------------------------------------------------------
SPIN_DAYS    <- 90                    # spin-up length before cal_start
cal_start    <- as.POSIXct(start, tz = "UTC")
cal_end      <- as.POSIXct(end, tz = "UTC")
sim_start    <- cal_start - SPIN_DAYS*24

obs_sim      <- obs[index(obs) >= sim_start & index(obs) <= cal_end, ]
obs_eval     <- obs[index(obs) >= cal_start & index(obs) <= cal_end, ]
obs_q        <- obs_eval[, "Discharge_m3_s"]   # keep as xts for time alignment

# -- Confirm outlet HRU --------------------------------------------------------
for(ii in seq_along(mdl$hru)){
  h <- mdl$hru[[ii]]
  if(length(h$sf_flow_direction$id) == 0 & length(h$sz_flow_direction$id) == 0)
    cat("Outlet HRU id:", h$id, "| is channel:", !is.na(h$class$endNode), "\n")
}
OUTLET_ID <- 0L

# =============================================================================
# 2. FLOW PRESENCE/ABSENCE OBSERVATIONS
# =============================================================================
#all_24 <- read_csv("/Users/johnmorgan/Documents/VT Research/HBTopModel/STIC_Synthesis/02_BinaryMatrix/Binary_hb_w3.csv")
all_24 <- read_csv("/Users/johnmorgan/Documents/VT Research/HBTopModel/STIC_Synthesis/all_24.csv") %>% 
  filter(wshed == "W3") %>% 
  select(datetime, wetdry, number, lat, long) %>% 
  filter(number %in% sensor_hru_key$sensor_id)



all_23 <- read_csv("/Users/johnmorgan/Documents/VT Research/HBTopModel/STIC_Synthesis/all_23.csv") %>% 
  filter(wshed == "W3") %>% 
  select(datetime, wetdry, ID, lat, long) %>% 
  rename(number = ID) %>% 
  filter(number %in% sensor_hru_key$sensor_id)



stic_locs <- all_23 %>% 
  bind_rows(all_24) %>%
  select(number, lat, long) %>%
  distinct() %>%
  vect(geom = c("long", "lat"), crs = "EPSG:4326") %>%
  project("EPSG:26919")

all_out <- do.call(rbind, lapply(seq_len(nrow(stic_locs)), function(i){
  dist_table <- distance(stic_locs[i,], vect_final)
  idx        <- which.min(dist_table)[1]
  data.frame(ID         = stic_locs[i,]$number,
             channel_ID = vect_final[idx,]$name,
             dist       = min(dist_table))
}))

sensor_hru_key
pa_obs <- 
  all_23 %>% 
  bind_rows(all_24) %>%
  filter(number %in% sensor_hru_key$sensor_id) %>%
  select(datetime, number, wetdry) %>%
  rename(ID = number) %>%
  left_join(sensor_hru_key, by = c("ID" = "sensor_id")) %>%
  #left_join(chosen, by = "ID", relationship = "many-to-one") %>%
  mutate(
    observed = as.integer(case_when(wetdry == "wet" ~ 1, wetdry == "dry" ~ 0)),
    datetime = as.POSIXct(datetime, tz = "UTC")
  ) %>%
  select(hru_id, datetime, observed) %>%
  filter(datetime >= cal_start, datetime <= cal_end,
         !is.na(hru_id), !is.na(observed))

cat("PA sites:", sort(unique(pa_obs$hru_id)), "\n")
cat("PA obs:", nrow(pa_obs), "| flowing:", sum(pa_obs$observed == 1),
    "| dry:", sum(pa_obs$observed == 0), "\n")

# Threshold for simulated "flowing" (m3/s). Adjust after Stage 1 diagnostics.
FLOW_THRESHOLD <- 1e-4   # 0.1 L/s is more defensible than 1e-8 for small streams

# =============================================================================
# 3. PARAMETER SPACE (Modified for 3 soil groups)
# =============================================================================
param_bounds <- data.frame(
  name  = c("t_0_1", "t_0_2", "t_0_3", 
            "m_1",   "m_2",   "m_3",   
            "t_d_1", "t_d_2", "t_d_3", 
            "s_rzmax", "c_sf", "T_T", "C_f"),
            
  # Lower bounds
  lower = c(0.5,   0.1,   0.01,   # t_0: Group 1 (high), Group 2 (mod), Group 3 (low)
            0.01,  0.02,  0.001,  # m:   Group 1 (mod), Group 2 (high/deep), Group 3 (low/shallow)
            3600,  43200, 3600,   # t_d: Group 1 (fast), Group 2 (slow), Group 3 (very fast)
            0.03,  0.01, -2.0,  0.1), # Other parameters (from original framework)
            
  # Upper bounds
  upper = c(5.0,   1.0,   0.5,    # t_0 
            0.03,  0.05,  0.01,   # m   
            28800, 172800,14400,  # t_d 
            0.3,   0.5,   2.0,  5.0), # Other parameters
            
  # Log-uniform sampling flags
  log   = c(TRUE,  TRUE,  TRUE,   # t_0 (usually spans orders of magnitude)
            FALSE, FALSE, FALSE,  # m (usually sampled uniformly within its narrow range)
            TRUE,  TRUE,  TRUE,   # t_d (spans orders of magnitude)
            FALSE, TRUE,  FALSE, TRUE),
            
  stringsAsFactors = FALSE
)

n_params <- nrow(param_bounds)

strip_obs <- function(obs_xts){
  drop <- c("Discharge_m3_s", "Discharge_ls", "Discharge")
  keep <- !(names(obs_xts) %in% drop)
  obs_xts[, keep]
}

# =============================================================================
# 4. MODEL RUNNER (Modified)
# =============================================================================
run_model <- function(params,
                      #mdl_template,
                      obs,
                      outlet_id  = 0L,
                      pa_hru_ids = sort(unique(pa_obs$hru_id)),
                      sub_step   = 3600,
                      include_sz = FALSE){        
  
  mdl_template <- readRDS( file.path(".", "7th_attempt", "dynaGISgoodChannel7","model_ready.rds"))
  hru <- mdl_template$hru

  obs_input  <- strip_obs(obs)
  total_area <- sum(sapply(hru, function(h) h$properties["area"]))
  
  q_init <- as.numeric(na.omit(obs$Discharge_m3_s))[1]
  if(!is.finite(q_init) || q_init <= 0) q_init <- 1e-5
  
  # --- CONSTANTS ---
  CONSTANT_V_CH <- 1.0  # Adjust your fixed channel velocity here
  #CONSTANT_N_CH <- 0.05 # Adjust your fixed Manning's n (if using kinematic)
  CONSTANT_C_WH <- 0.2  # Adjust your fixed water holding capacity
  
  for(ii in seq_along(hru)){
    is_channel <- isTRUE(!is.na(hru[[ii]]$class$endNode))
    
    if(!is_channel){
      # Fetch the soil index (1, 2, or 3) from the map generated earlier
      s_idx <- hru_soil_map$soil_idx[hru_soil_map$hru_index == ii]
      
      hru[[ii]]$sz$parameters["t_0"]     <- params[paste0("t_0_", s_idx)]
      hru[[ii]]$sz$parameters["m"]       <- params[paste0("m_", s_idx)]
      hru[[ii]]$uz$parameters["t_d"]     <- params[paste0("t_d_", s_idx)]
      hru[[ii]]$rz$parameters["s_rzmax"] <- params["s_rzmax"]
      
      if(hru[[ii]]$sf$type == "cnst"){
        hru[[ii]]$sf$parameters["c_sf"]  <- params["c_sf"]
      }
    #} else {
      #hru[[ii]]$sz$parameters["t_0"]     <- 0.001
      #hru[[ii]]$sz$parameters["m"]       <- 0.001        
      #hru[[ii]]$rz$parameters["s_rzmax"] <- 0.001
      
      # Apply Constants
      if(hru[[ii]]$sf$type == "cnst"){
        hru[[ii]]$sf$parameters["c_sf"]  <- CONSTANT_V_CH
      } else if(hru[[ii]]$sf$type == "kin"){
        hru[[ii]]$sf$parameters["n"]     <- CONSTANT_N_CH
        hru[[ii]]$sf$parameters["t_raf"] <- 999.9
        hru[[ii]]$sf$parameters["s_raf"] <- 0
      }
    }
    
    # Snow parameters (C_wh is constant)
    hru[[ii]]$snow$parameters["T_T"]  <- params["T_T"]
    hru[[ii]]$snow$parameters["C_f"]  <- params["C_f"]
    hru[[ii]]$snow$parameters["C_wh"] <- CONSTANT_C_WH
    
    # Initial conditions
    s_rzmax_val <- if(!is_channel) as.numeric(params["s_rzmax"]) else 0.001
    hru[[ii]]$initialisation["s_rz_0"]    <- 0.7
    hru[[ii]]$initialisation["r_uz_sz_0"] <- q_init / total_area
    
    # if("s_sz_0" %in% names(hru[[ii]]$initialisation)){
    #   hru[[ii]]$initialisation["s_sz_0"] <- 0.001
    # }
  }
  
  pa_ids <- unique(pa_hru_ids)
  out_rows <- data.frame(
    name  = c("q_out", if(include_sz) "q_out_sz", paste0("q_sf_hru", pa_ids)),
    id    = c(outlet_id, if(include_sz) outlet_id, pa_ids),
    flux  = c("q_sf", if(include_sz) "q_sz", rep("q_sf", length(pa_ids))),
    scale = 1.0,
    stringsAsFactors = FALSE
  )
  
  result <- tryCatch({
    cm <- dynatop$new(hru, map = mdl_template$map)
    cm$add_data(obs_input)
    cm$initialise()
    cm$sim(out_rows, sub_step = sub_step)$get_output()
  }, error = function(e){
    warning("run_model error: ", conditionMessage(e))
    NULL
  })
  return(result)
}

default_params <- c("t_0_1" = 0.135, "t_0_2"= 0.135, "t_0_3"= 0.135, 
            "m_1"=0.040,   "m_2"=0.040,   "m_3"=0.040,   
            "t_d_1"=28800, "t_d_2"=28800, "t_d_3"=28800, 
            "s_rzmax"=0.1, "c_sf"=0.3, "T_T"=0.0, "C_f"=0.002)
run_model(default_params,
          obs)
# =============================================================================
# 5. DIAGNOSTIC (revised)
# =============================================================================
run_diagnostic <- function(mdl, obs_sim, obs_eval, pa_obs,
                           outlet_id = 0L, sub_step = 900, include_sz = FALSE){
  
  obs_input <- strip_obs(obs_sim)
  nms <- c("q_sf_out", "s_sf_out", "s_sz_out", "q_sz_out")
  if(include_sz) nms <- c(nms, "q_total_out")
  
  diag_rows <- data.frame(
    name  = c("q_sf_out", "s_sf_out", "s_sz_out", "q_sz_out"),
    id    = outlet_id,
    flux  = c("q_sf", "s_sf", "s_sz", "q_sz"),
    scale = 1.0,
    stringsAsFactors = FALSE
  )
  
  cm <- dynatop$new(mdl$hru, map = mdl$map)
  cm$add_data(obs_input)
  cm$initialise()
  diag_out <- cm$sim(diag_rows, sub_step = sub_step)$get_output()
  
  # Evaluate only over cal window
  eval_idx <- index(diag_out) >= index(obs_eval)[1] & index(diag_out) <= tail(index(obs_eval),1)
  q_sf_win <- diag_out[eval_idx, "q_sf_out"]
  q_sz_win <- diag_out[eval_idx, "q_sz_out"]
  
  q_tot <- if(include_sz) q_sf_win + q_sz_win else q_sf_win
  
  cat("=== Diagnostics (evaluation window) ===\n")
  cat("q_sf range (m3/s):", signif(range(as.numeric(q_sf_win)), 3), "\n")
  if(include_sz) cat("q_sz range (m3/s):", signif(range(as.numeric(q_sz_win)), 3), "\n")
  cat("Sim Total range  :", signif(range(as.numeric(q_tot)), 3), "\n")
  cat("obs_q range      :", signif(range(as.numeric(obs_eval$Discharge_m3_s), na.rm=TRUE), 3), "\n")
  cat("Bias (sim/obs)   :", signif(mean(q_tot, na.rm=TRUE)/mean(obs_eval$Discharge_m3_s, na.rm=TRUE), 3), "\n")
  
  s_sf_vals <- as.numeric(diag_out[,"s_sf_out"])
  if(tail(s_sf_vals,1) > 10 * s_sf_vals[1] + 1){
    cat("WARNING: s_sf growing unboundedly.\n")
  } else {
    cat("s_sf appears stable.\n")
  }
  
  # PA quick look
  pa_ids <- sort(unique(pa_obs$hru_id))
  pa_rows <- data.frame(
    name  = paste0("q_sf_hru", pa_ids),
    id    = pa_ids,
    flux  = "q_sf",
    scale = 1.0,
    stringsAsFactors = FALSE
  )
  cm2 <- dynatop$new(mdl$hru, map = mdl$map)
  cm2$add_data(obs_input)
  cm2$initialise()
  pa_out <- cm2$sim(pa_rows, sub_step = sub_step)$get_output()
  
  cat("\n=== PA site q_sf (during eval window) ===\n")
  for(hid in pa_ids){
    col <- paste0("q_sf_hru", hid)
    sub <- pa_obs[pa_obs$hru_id == hid & pa_obs$observed == 1, ]
    if(nrow(sub) == 0 || !(col %in% names(pa_out))) next
    win <- pa_out[index(pa_out) >= cal_start & index(pa_out) <= cal_end, col]
    cat(sprintf("HRU %d | min=%.2e max=%.2e (obs flow n=%d)\n",
                hid, min(win, na.rm=TRUE), max(win, na.rm=TRUE), nrow(sub)))
  }
  invisible(list(diag_out = diag_out, pa_out = pa_out))
}

cat("Running diagnostic with template defaults (include_sz = TRUE)...\n")
Sys.time()
diag <- run_diagnostic(mdl, obs_sim, obs_eval, pa_obs,
                       outlet_id = OUTLET_ID, include_sz = TRUE)
Sys.time()
# After seeing diagnostics, set this to TRUE if q_sz is a major part of discharge
INCLUDE_SZ_AT_OUTLET <- FALSE

# =============================================================================
# 6. OBJECTIVE FUNCTIONS
# =============================================================================
kge <- function(sim, obs){
  both  <- merge(sim, obs, join = "inner")
  s     <- as.numeric(both[, 1])
  o     <- as.numeric(both[, 2])
  valid <- is.finite(s) & is.finite(o) & abs(s) < 1e4
  if(sum(valid) < 10) return(-Inf)
  s <- s[valid]; o <- o[valid]
  r <- cor(s, o); if(!is.finite(r)) return(-Inf)
  a <- sd(s) / sd(o)
  b <- mean(s) / mean(o)
  1 - sqrt((r - 1)^2 + (a - 1)^2 + (b - 1)^2)
}

bias_ratio <- function(sim, obs){
  both <- merge(sim, obs, join = "inner")
  mean(as.numeric(both[,1]), na.rm=TRUE) / mean(as.numeric(both[,2]), na.rm=TRUE)
}

csi_presence_absence <- function(sim_output, pa_obs, threshold = FLOW_THRESHOLD){
  TP <- FP <- FN <- TN <- 0L
  for(hid in unique(pa_obs$hru_id)){
    col_name <- paste0("q_sf_hru", hid)
    if(!col_name %in% names(sim_output)) next
    sub_obs <- pa_obs[pa_obs$hru_id == hid, ]
    if(nrow(sub_obs) == 0) next
    sub_xts <- xts(sub_obs$observed, order.by = as.POSIXct(sub_obs$datetime, tz = "UTC"))
    both    <- merge(sim_output[, col_name], sub_xts, join = "inner")
    if(nrow(both) == 0) next
    sim_bin <- as.integer(as.numeric(both[, 1]) >= threshold)
    obs_bin <- as.integer(as.numeric(both[, 2]))
    valid   <- !is.na(sim_bin) & !is.na(obs_bin)
    sim_bin <- sim_bin[valid]; obs_bin <- obs_bin[valid]
    if(length(sim_bin) == 0) next
    TP <- TP + sum(sim_bin == 1L & obs_bin == 1L)
    FP <- FP + sum(sim_bin == 1L & obs_bin == 0L)
    FN <- FN + sum(sim_bin == 0L & obs_bin == 1L)
    TN <- TN + sum(sim_bin == 0L & obs_bin == 0L)
  }
  denom <- TP + FP + FN
  if(denom == 0) return(NULL)
  list(
    CSI = TP / denom,
    #CSI = if((TP+FP+FN+TN)>0) (TP+TN)/(TP+FP+FN+TN) else 0,
    POD = if((TP+FN)>0) TP/(TP+FN) else 0,
    FAR = if((TP+FP)>0) FP/(TP+FP) else 0,
    F1  = if((2*TP+FP+FN)>0) 2*TP/(2*TP+FP+FN) else 0,
    accuracy = if((TP+FP+FN+TN)>0) (TP+TN)/(TP+FP+FN+TN) else 0,
    TP=TP, FP=FP, FN=FN, TN=TN
  )
}

# =============================================================================
# 7. GLUE INFRASTRUCTURE
# =============================================================================
W_KGE <- 0.5
W_CSI <- 0.5
CSI_THRESHOLD <- 0.0

glue_likelihood <- function(kge_val, csi_val, kge_thresh){
  if(!is.finite(kge_val) || kge_val < kge_thresh) return(0)
  if(!is.finite(csi_val) || csi_val < CSI_THRESHOLD) return(0)
  kge_scaled <- (kge_val - kge_thresh) / (1 - kge_thresh)
  csi_scaled <- csi_val
  L_kge <- if(W_KGE > 0) kge_scaled ^ W_KGE else 1
  L_csi <- if(W_CSI > 0) csi_scaled ^ W_CSI else 1
  L_kge * L_csi
}

# Vectorized parameter sampler
generate_lhs <- function(n_samp, pb){
  d <- randomLHS(n_samp, nrow(pb))
  M <- matrix(NA, nrow=n_samp, ncol=nrow(pb))
  colnames(M) <- pb$name
  for(i in seq_len(nrow(pb))){
    lo <- pb$lower[i]; up <- pb$upper[i]
    if(pb$log[i]){
      loglo <- log10(lo); logup <- log10(up)
      M[,i] <- 10^(d[,i]*(logup - loglo) + loglo)
    } else {
      M[,i] <- d[,i]*(up - lo) + lo
    }
  }
  M
}

# Generic GLUE evaluator
glue_eval <- function(param_vec,
                      include_sz = FALSE,
                      use_pa     = TRUE,
                      kge_thresh = 0.0){
  
  params <- setNames(param_vec, param_bounds$name)
  
  # Hard guards for multiple soil groups
  if(any(params[c("t_d_1", "t_d_2", "t_d_3")] < 900)) return(list(L=0, kge=NA, csi=NA, bias=NA))
  if(params["s_rzmax"] <= 0)  return(list(L=0, kge=NA, csi=NA, bias=NA))
  
  sim <- run_model(params, #mdl_template, 
                   obs_sim,
                   outlet_id  = OUTLET_ID,
                   pa_hru_ids = sort(unique(pa_obs$hru_id)),
                   include_sz = include_sz)
  
  if(is.null(sim)) return(list(L=0, kge=NA, csi=NA, bias=NA))
  
  # Outlet discharge (xts, full length incl. spinup; merge with obs_q auto-subsets)
  sim_q <- if(include_sz && "q_out_sz" %in% names(sim)){
    sim[,"q_out"] + sim[,"q_out_sz"]
  } else {
    sim[,"q_out"]
  }
  
  if(any(!is.finite(sim_q))) return(list(L=0, kge=NA, csi=NA, bias=NA))
  
  kge_val   <- kge(sim_q, obs_q)
  bias_val  <- bias_ratio(sim_q, obs_q)
  pa_scores <- if(use_pa) csi_presence_absence(sim, pa_obs) else NULL
  csi_val   <- if(!is.null(pa_scores) && is.finite(pa_scores$CSI)) pa_scores$CSI else 0
  #accuracy <- if(!is.null(pa_scores) && is.finite(pa_scores$accuracy)) pa_scores$accuracy else 0
  
  # Inside glue_eval, replace the single L <- ... line with:
  if(!use_pa){
    # Screening: pure KGE ranking, ignore CSI
    if(!is.finite(kge_val) || kge_val < kge_thresh){
      L <- 0
    } else {
      L <- ((kge_val - kge_thresh) / (1 - kge_thresh)) ^ W_KGE
    }
  } else {
    L <- glue_likelihood(kge_val, csi_val, kge_thresh)
  }
  list(L = L, kge = kge_val, csi = csi_val, bias = bias_val)
}

run_glue <- function(n_samples,
                     param_bounds,
                     #mdl_template,
                     obs_sim,
                     obs_q,
                     pa_obs,
                     outlet_id    = 0L,
                     n_cores      = 1L,
                     seed         = 42L,
                     use_pa       = TRUE,
                     kge_thresh   = 0.0,
                     include_sz   = FALSE,
                     sub_step     = 3600){
  
  set.seed(seed)
  n_params <- nrow(param_bounds)
  param_matrix <- generate_lhs(n_samples, param_bounds)
  
  cat(sprintf("Running %d GLUE samples (%s) on %d core(s)...\n",
              n_samples,
              if(use_pa) "KGE+CSI" else "KGE only",
              n_cores))
  
  # Local wrapper so that parApply/exporting is clean
  obj_wrap <- function(v){
    glue_eval(v,
              include_sz = include_sz,
              use_pa     = use_pa,
              kge_thresh = kge_thresh)
  }
  
  if(n_cores > 1L){
    cl <- makeCluster(n_cores)
    on.exit(stopCluster(cl), add = TRUE)
    clusterExport(cl,
      c(#"mdl_template",
        "obs_sim","obs_q","pa_obs","param_bounds",
        "OUTLET_ID","FLOW_THRESHOLD","W_KGE","W_CSI","CSI_THRESHOLD",
        "run_model","strip_obs","kge","bias_ratio","csi_presence_absence",
        "glue_likelihood","glue_eval","generate_lhs", "hru_soil_map"),
      envir = environment())
    clusterEvalQ(cl, { library(dynatop); library(xts) })
    score_list <- parApply(cl, param_matrix, 1, obj_wrap)
  } else {
    score_list <- apply(param_matrix, 1, obj_wrap)
  }
  
  results <- data.frame(
    param_matrix,
    L    = sapply(score_list, `[[`, "L"),
    kge  = sapply(score_list, `[[`, "kge"),
    csi  = sapply(score_list, `[[`, "csi"),
    bias = sapply(score_list, `[[`, "bias")
  )
  
  n_behav <- sum(results$L > 0, na.rm = TRUE)
  cat(sprintf("Behavioural: %d / %d (%.1f%%) | Best KGE: %.3f\n",
              n_behav, n_samples, 100*n_behav/n_samples,
              max(results$kge, na.rm=TRUE)))
  
  if(n_behav == 0)
    cat("No behavioural runs — try lowering kge_thresh or check run_model.\n")
  
  results$L_norm <- if(sum(results$L, na.rm=TRUE) > 0){
    results$L / sum(results$L, na.rm=TRUE)
  } else NA_real_
  
  results[order(-results$L), ]
}

# =============================================================================
# 8. UNCERTAINTY BOUNDS (unchanged logic, accepts include_sz via run_model)
# =============================================================================
glue_uncertainty_bounds <- function(glue_results, param_bounds,
                                    #mdl_template, 
                                    obs_sim, obs_q,
                                    outlet_id = 0L,
                                    quantiles = c(0.05, 0.95),
                                    include_sz = FALSE){
  
  behav <- glue_results[glue_results$L > 0 & !is.na(glue_results$L), ]
  if(nrow(behav) == 0) stop("No behavioural runs.")
  cat(sprintf("Bounds from %d behavioural runs...\n", nrow(behav)))
  
  sim_list <- lapply(seq_len(nrow(behav)), function(i){
    p <- setNames(as.numeric(behav[i, param_bounds$name]), param_bounds$name)
    s <- run_model(p, mdl_template, obs_sim, outlet_id,
                   include_sz = include_sz)
    if(is.null(s)) return(NULL)
    q <- if(include_sz && "q_out_sz" %in% names(s)) s[,"q_out"] + s[,"q_out_sz"] else s[,"q_out"]
    as.numeric(q)
  })
  
  good  <- !sapply(sim_list, is.null)
  sims  <- sim_list[good]
  L_w   <- behav$L_norm[good]
  L_w   <- L_w / sum(L_w)
  
  sim_mat <- do.call(cbind, sims)
  # align with obs_q (just in case spinup mismatched length)
  nobs <- length(na.omit(as.numeric(obs_q)))
  if(nrow(sim_mat) != nobs){
    # attempt to trim to evaluation window index length
    sim_mat <- sim_mat[(nrow(sim_mat)-nobs+1):nrow(sim_mat), , drop=FALSE]
  }
  
  bounds <- data.frame(
    datetime = index(obs_q),
    q_obs    = as.numeric(obs_q),
    q_best   = sim_mat[, which.max(L_w)],
    q_mean   = apply(sim_mat, 1, weighted.mean, w = L_w)
  )
  for(q in quantiles){
    col <- paste0("q_", sprintf("%02d", round(q*100)))
    bounds[[col]] <- apply(sim_mat, 1, function(row){
      ord <- order(row); cw <- cumsum(L_w[ord])
      row[ord][which(cw >= q)[1]]
    })
  }
  bounds
}

# Plotting functions (keep originals, just pass result)
plot_glue_bounds <- function(bounds){
  t_idx <- bounds$datetime
  y_max <- max(c(bounds$q_obs, bounds$q_95), na.rm=TRUE)*1.1
  plot(t_idx, bounds$q_obs, type="l", col="black", lwd=1.5,
       ylim=c(0, y_max), main="GLUE bounds vs observed [m3/s]", xlab="", ylab="m3/s")
  polygon(c(t_idx, rev(t_idx)), c(bounds$q_05, rev(bounds$q_95)),
          col=adjustcolor("steelblue", alpha.f=0.3), border=NA)
  lines(t_idx, bounds$q_best, col="firebrick", lwd=1.5)
  lines(t_idx, bounds$q_mean, col="steelblue", lwd=1, lty=2)
  legend("topright", legend=c("Obs","Best","Mean","5-95%"),
         col=c("black","firebrick","steelblue",adjustcolor("steelblue",0.4)),
         lwd=c(1.5,1.5,1,8), lty=c(1,1,2,1))
}

plot_glue_dotty <- function(glue_results, param_bounds){
  behav    <- glue_results[glue_results$L > 0 & !is.na(glue_results$L), ]
  nonbehav <- glue_results[is.na(glue_results$L) | glue_results$L == 0, ]
  n_plots  <- nrow(param_bounds)
  nr <- ceiling(sqrt(n_plots)); nc <- ceiling(n_plots/nr)
  par(mfrow=c(nr,nc))
  for(pn in param_bounds$name){
    y_max <- if(nrow(behav)>0) max(behav$L, na.rm=TRUE) else 1
    xlim  <- c(param_bounds$lower[param_bounds$name==pn],
               param_bounds$upper[param_bounds$name==pn])
    if(nrow(nonbehav)>0)
      plot(nonbehav[[pn]], nonbehav$L, xlab=pn, ylab="Likelihood",
           pch=16, cex=0.4, col="grey70", ylim=c(0.3,y_max), xlim=xlim)
    else
      plot(0,0, xlab=pn, ylab="Likelihood", ylim=c(0.3,y_max), xlim=xlim, type="n")
    if(nrow(behav)>0)
      points(behav[[pn]], behav$L, pch=16, cex=0.7, col="firebrick")
    abline(h=0, lty=2, col="grey50")
  }
  par(mfrow=c(1,1))
}

# Helper: tighten bounds from behavioral percentile + margin
tighten_bounds <- function(pb, runs, prob = c(0.02, 0.98), pad = 1.2){
  pb2 <- pb
  for(i in seq_len(nrow(pb))){
    nm <- pb$name[i]
    vals <- runs[[nm]]
    if(length(vals) < 5) next
    q <- quantile(vals, probs = prob, na.rm = TRUE)
    if(pb$log[i]){
      lo <- q[1] / pad
      up <- q[2] * pad
    } else {
      w  <- (q[2] - q[1]) * (pad - 1)/2
      lo <- q[1] - w
      up <- q[2] + w
    }
    pb2$lower[i] <- max(pb2$lower[i], lo)
    pb2$upper[i] <- min(pb2$upper[i], up)
    if(pb2$log[i]) pb2$lower[i] <- max(pb2$lower[i], 1e-7)
  }
  pb2
}

# =============================================================================
# 9. RUN — Two-stage GLUE
# =============================================================================


# -- Stage 1: Wide screen on discharge only -----------------------------------
cat("\n============= STAGE 1: DISCHARGE SCREENING =============\n")
Sys.time()
screen_results <- run_glue(
  n_samples    = 1,          # 3k cheap runs to find the sweet spot
  param_bounds = param_bounds,
  #mdl_template = mdl,
  obs_sim      = obs_sim,
  obs_q        = obs_q,
  pa_obs       = pa_obs,
  outlet_id    = OUTLET_ID,
  n_cores      = 1L,
  seed         = 1L,
  use_pa       = FALSE,         # ignore PA for now
  kge_thresh   = 0.0,          # permissive but not garbage
  include_sz   = FALSE
)
# for(z in 1:48){
# screen_results <- run_glue(
#   n_samples    = 1000,          # 3k cheap runs to find the sweet spot
#   param_bounds = param_bounds,
#   mdl_template = mdl,
#   obs_sim      = obs_sim,
#   obs_q        = obs_q,
#   pa_obs       = pa_obs,
#   outlet_id    = OUTLET_ID,
#   n_cores      = 7L,
#   seed         = 236L+z,
#   use_pa       = FALSE,         # ignore PA for now
#   kge_thresh   = -0.4,          # permissive but not garbage
#   include_sz   = FALSE
# )
# #Sys.time()
# saveRDS(screen_results, 
#         paste0("./glue_results_soils/glue_screen1000_seed",236+z,".rds"))
# print(paste0("run ",z," out of 48"))
#         
# }

## 5/10/26, did about 30,000 runs, no kge above 0
## tighten bounds, and do again
files <- list.files("./glue_results_soils")
for(i in 1:length(files)){
  file <- files[i]
  
  raw_glue <- readRDS(paste0("./glue_results_soils/",file))
  screen_glue <- raw_glue[raw_glue$L > 0 & !is.na(raw_glue$L), ]
  
  if(i == 1) alldat <- screen_glue
  if(i > 1) alldat <- rbind(alldat, screen_glue)
}
screen_behav <- alldat %>% 
  filter(kge >= -0.4)
screen_behav[1:10,]
screen_behav %>% 
  tibble() %>% 
arrange(desc(kge))
# Tighten to 5th–95th percentile of behavioral sets, with a little padding
pb_focused <- tighten_bounds(param_bounds, screen_behav, prob = c(0.05, 0.95), pad = 1.2)


screen_results <- run_glue(
  n_samples    = 10,          # 3k cheap runs to find the sweet spot
  param_bounds = best_params,
  mdl_template = mdl,
  obs_sim      = obs_sim,
  obs_q        = obs_q,
  pa_obs       = pa_obs,
  outlet_id    = OUTLET_ID,
  n_cores      = 6L,
  seed         = 301L,
  use_pa       = FALSE,         # ignore PA for now
  kge_thresh   = 0.0,          # permissive but not garbage
  include_sz   = FALSE
)
#Sys.time()
saveRDS(screen_results, 
        paste0("./glue_results_soils/glue_screen1000_seed300.rds"))

cat("\nTop 10 Stage-2 behavioral runs:\n")
best <- screen_results %>% 
  tibble() %>% 
arrange(desc(kge))

best_params <- setNames(as.numeric(best[1, param_bounds$name]), param_bounds$name)
best_sim <- run_model(best_params, mdl, obs_sim,
                      outlet_id  = OUTLET_ID,
                      pa_hru_ids = sort(unique(pa_obs$hru_id)),
                      include_sz = FALSE)
plot(best_sim$q_out)

######## second round
# Extract winner
winner <- screen_results[which.max(screen_results$kge), ]
winner_params <- setNames(as.numeric(winner[, param_bounds$name]), param_bounds$name)
best_params
# Focused bounds around the sweet spot (±1 order of magnitude for logs, ±2× for linear)
param_bounds_focused <- data.frame(
  name  = c("t_0","m","t_d","s_rzmax","c_sf","n_ch","v_ch","T_T","C_f","C_wh"),
  lower = c(0.03,   0.001,  1800,   0.03,   0.05,  0.01,  0.3,  -1.0,  0.5,   0.05),
  upper = c(5.0,    0.05,   86400,  0.3,    1.5,   0.1,   3.0,   2.0,  10.0,  0.6),
  log   = c(TRUE,   FALSE,  TRUE,   FALSE,  TRUE,  FALSE, FALSE, FALSE, TRUE,  FALSE),
  stringsAsFactors = FALSE
)
files <- list.files("./glue_results")
for(i in 1:length(files)){
  file <- files[i]
  
  raw_glue <- readRDS(paste0("./glue_results/",file))
  screen_glue <- raw_glue[raw_glue$L > 0 & !is.na(raw_glue$L), ]
  
  if(i == 1) alldat <- screen_glue
  if(i > 1) alldat <- rbind(alldat, screen_glue)
}
screen_behav <- alldat

# Tighten to 5th–95th percentile of behavioral sets, with a little padding
pb_focused <- tighten_bounds(param_bounds, screen_behav, prob = c(0.05, 0.95), pad = 1.2)
cat("\nFocused bounds after Stage 1:\n")
print(pb_focused)

# -- Stage 2: Focused GLUE with PA --------------------------------------------
cat("\n============= STAGE 2: FOCUSED GLUE (KGE + PA) =============\n")
Sys.time()
glue_results <- run_glue(
  n_samples    = 300,          # denser sample in reduced space
  param_bounds = pb_focused,
  mdl_template = mdl,
  obs_sim      = obs_sim,
  obs_q        = obs_q,
  pa_obs       = pa_obs,
  outlet_id    = OUTLET_ID,
  n_cores      = 7L,
  seed         = 107L,
  use_pa       = TRUE,
  kge_thresh   = 0.0,           # must beat the mean predictor
  include_sz   = FALSE
)
saveRDS(glue_results, "./glue_PA_acc_results/glue_screen3000_seed104_PA.rds")
tk <- readRDS("./glue_PA_acc_results/glue_screen3000_seed104_PA.rds")
View(tk)
best <- glue_results %>% 
  tibble() %>% 
arrange(desc(kge))
best
index <- which(glue_results$kge == max(glue_results$kge, na.rm = TRUE))
glue_results[which(glue_results$kge == max(glue_results$kge, na.rm = TRUE)),]
#big_best <- best
# =============================================================================
# 10. EVALUATE BEST PARAMETER SET
# =============================================================================
cat("\nTop 10 Stage-2 behavioral runs:\n")
print(head(glue_results[glue_results$L > 0 & !is.na(glue_results$L),
                        c(param_bounds$name, "L", "kge", "csi", "bias")], 10))

best_params <- setNames(as.numeric(glue_results[index, pb_focused$name]), pb_focused$name)
best_sim <- run_model(best_params, mdl, obs_sim,
                      outlet_id  = OUTLET_ID,
                      pa_hru_ids = sort(unique(pa_obs$hru_id)),
                      include_sz = FALSE)

# Compose total discharge for plotting
best_q <- best_sim[,"q_out"]


cat("\nBest run evaluation:\n")
cat("KGE :", kge(best_q, obs_q), "\n")
cat("Bias:", bias_ratio(best_q, obs_q), "\n")

pa_best <- csi_presence_absence(best_sim, pa_obs)
if(!is.null(pa_best)){
  cat(sprintf("CSI=%.3f POD=%.3f FAR=%.3f TP=%d FP=%d FN=%d TN=%d\n ACC=%.3f",
              pa_best$CSI, pa_best$POD, pa_best$FAR,
              pa_best$TP, pa_best$FP, pa_best$FN, pa_best$TN, pa_best$accuracy))
}

# -- Discharge plot --
plot(merge(obs_q, best_q),
     main = "Observed vs Simulated discharge [m3/s]",
     legend.loc = "topright",
     ylim = c(0, max(c(as.numeric(obs_q), as.numeric(best_q)), na.rm=TRUE)*1.1),
     col = c("black", "firebrick"))
plot(merge(best_q, obs_q),
     main = "Observed vs Simulated discharge [m3/s]",
     legend.loc = "topright",
     ylim = c(0, max(c(as.numeric(obs_q), as.numeric(best_q)), na.rm=TRUE)*1.1),
     col = c("firebrick", "black"))

# -- Dotty plots for focused run --
#par(mar=c(3,3,3,3))

plot_glue_dotty(tk, pb_focused)

# -- PA site plots --
for(hid in sort(unique(pa_obs$hru_id))){
  col_name <- paste0("q_sf_hru", hid)
  if(!col_name %in% names(best_sim)) next
  sim_q <- best_sim[, col_name]
  sub   <- pa_obs[pa_obs$hru_id == hid, ]
  
  t_sim <- index(sim_q)
  ylims <- c(0, max(c(as.numeric(sim_q), FLOW_THRESHOLD*3), na.rm=TRUE))
  xlims <- c(min(pa_obs$datetime), max(all_23$datetime))
  
  plot(t_sim, as.numeric(sim_q), type="l",
       main = paste("HRU", hid, "— simulated q_sf vs observations 2023"),
       ylab = "m3/s", xlab = "", col = "steelblue", ylim = ylims
       ,xlim = xlims
       )
  abline(h = FLOW_THRESHOLD, lty = 2, col = "grey40")
  
  flowing <- sub[sub$observed == 1, ]
  dry     <- sub[sub$observed == 0, ]
  
  if(nrow(flowing) > 0){
    t_f <- sort(unique(as.POSIXct(flowing$datetime, tz="UTC")))
    rect(xleft = head(t_f,-1), xright = tail(t_f,-1),
         ybottom = 0, ytop = par("usr")[4],
         col = adjustcolor("green", 0.15), border = NA)
  }
  if(nrow(dry) > 0){
    t_d <- sort(unique(as.POSIXct(dry$datetime, tz="UTC")))
    rect(xleft = head(t_d,-1), xright = tail(t_d,-1),
         ybottom = 0, ytop = par("usr")[4],
         col = adjustcolor("red", 0.1), border = NA)
  }
  legend("topright",
         legend = c("Simulated q_sf", "Flow threshold", "Obs: flowing", "Obs: dry"),
         col = c("steelblue","grey40", adjustcolor("green",0.4), adjustcolor("red",0.3)),
         lty = c(1,2,NA,NA), pch = c(NA,NA,15,15), cex=0.8)
}

## SUMMER 2024
for(hid in sort(unique(pa_obs$hru_id))){
  col_name <- paste0("q_sf_hru", hid)
  if(!col_name %in% names(best_sim)) next
  sim_q <- best_sim[, col_name]
  sub   <- pa_obs[pa_obs$hru_id == hid, ]
  
  t_sim <- index(sim_q)
  ylims <- c(0, max(c(as.numeric(sim_q), FLOW_THRESHOLD*3), na.rm=TRUE))
  xlims <- c(min(all_24$datetime), max(all_24$datetime))
  
  plot(t_sim, as.numeric(sim_q), type="l",
       main = paste("HRU", hid, "— simulated q_sf vs observations 2024"),
       ylab = "m3/s", xlab = "", col = "steelblue", ylim = ylims
       ,xlim = xlims
       )
  abline(h = FLOW_THRESHOLD, lty = 2, col = "grey40")
  
  flowing <- sub[sub$observed == 1, ]
  dry     <- sub[sub$observed == 0, ]
  
  if(nrow(flowing) > 0){
    t_f <- sort(unique(as.POSIXct(flowing$datetime, tz="UTC")))
    rect(xleft = head(t_f,-1), xright = tail(t_f,-1),
         ybottom = 0, ytop = par("usr")[4],
         col = adjustcolor("green", 0.15), border = NA)
  }
  if(nrow(dry) > 0){
    t_d <- sort(unique(as.POSIXct(dry$datetime, tz="UTC")))
    rect(xleft = head(t_d,-1), xright = tail(t_d,-1),
         ybottom = 0, ytop = par("usr")[4],
         col = adjustcolor("red", 0.1), border = NA)
  }
  legend("topright",
         legend = c("Simulated q_sf", "Flow threshold", "Obs: flowing", "Obs: dry"),
         col = c("steelblue","grey40", adjustcolor("green",0.4), adjustcolor("red",0.3)),
         lty = c(1,2,NA,NA), pch = c(NA,NA,15,15), cex=0.8)
}




# -- Uncertainty bounds --
n_behav <- sum(glue_results$L > 0, na.rm=TRUE)
#if(n_behav >= 10){
  bounds <- glue_uncertainty_bounds(best[1:10,], pb_focused,
                                    mdl, obs_sim, obs_q,
                                    outlet_id = OUTLET_ID,
                                    include_sz = FALSE)
  plot_glue_bounds(bounds)

```


```{r}
# =============================================================================
# dynatop Parameterisation — simplified and efficient
# Calibrates against: NSE, KGE, CSI, Accuracy (presence/absence)
# Soil-group-specific: t_0, m, t_d  (h$class$soil = 1 / 2 / 3)
# Shared across all HRUs: s_rzmax, c_sf, T_T, C_f
# Model reloaded from disk on every run — no state corruption possible
# =============================================================================

# =============================================================================
# 1.  PATHS
# =============================================================================
MDL_PATH <- file.path(".", "7th_attempt", "dynaGISgoodChannel7", "model_ready.rds")
OBS_PATH <- file.path(".", "7th_attempt", "obs.rds")
OUT_DIR  <- file.path(".", "7th_attempt", "glue_results")
dir.create(OUT_DIR, showWarnings = FALSE)

# =============================================================================
# 2.  DATA
# =============================================================================
obs_full <- readRDS(OBS_PATH)

cal_start <- as.POSIXct("2021-01-01", tz = "UTC")
cal_end   <- as.POSIXct("2024-11-26", tz = "UTC")
spin_days <- 180L
sim_start <- cal_start - spin_days * 86400

obs_sim <- obs_full[index(obs_full) >= sim_start & index(obs_full) <= cal_end, ]
obs_q   <- obs_full[index(obs_full) >= cal_start & index(obs_full) <= cal_end,
                    "Discharge_m3_s"]

# pa_obs: data frame with columns hru_id (int), datetime (POSIXct UTC), observed (0/1)
# pa_obs <- readRDS("pa_obs.rds")   # load yours here
pa_obs <- read_csv("./7th_attempt/pa_obs.csv") %>% 
  mutate(mins = lubridate::minute(datetime)) %>% 
  filter(mins == 0) %>% 
  select(!mins) %>% 
  mutate(datetime = as.POSIXct(datetime))
FLOW_THRESHOLD <- 1e-4
OUTLET_ID      <- 0L

strip_obs <- function(x){
  x[, !names(x) %in% c("Discharge_m3_s", "Discharge_ls", "Discharge")]
}

build_out_rows <- function(pa_hru_ids, outlet_id = OUTLET_ID){
  rbind(
    data.frame(name="q_out", id=outlet_id, flux="q_sf",
               scale=1.0, stringsAsFactors=FALSE),
    data.frame(name=paste0("q_sf_hru", pa_hru_ids),
               id=pa_hru_ids, flux="q_sf",
               scale=1.0, stringsAsFactors=FALSE)
  )
}

# =============================================================================
# 3.  PARAMETER BOUNDS
# =============================================================================
# Soil group meanings (h$class$soil):
#   1 = organic-rich near-stream  — high T0, moderate m, fast t_d
#   2 = deep hillslope mineral    — moderate T0, LOW m, slow t_d
#   3 = shallow rocky             — moderate-high T0, HIGH m, fast t_d

param_bounds <- data.frame(
  name  = c("t_0_1","t_0_2","t_0_3",
            "m_1",  "m_2",  "m_3",
            "t_d_1","t_d_2","t_d_3",
            "s_rzmax","c_sf","T_T","C_f"),
  lower = c(0.10,  0.01,  0.05,
            0.010, 0.001, 0.020,
            3600,  86400, 3600,
            0.03,  0.01, -2.0, 0.1),
  upper = c(5.00,  0.50,  2.00,
            0.040, 0.015, 0.060,
            28800, 259200, 14400,
            0.40,  0.50,  2.0,  5.0),
  log   = c(TRUE,  TRUE,  TRUE,
            FALSE, FALSE, FALSE,
            TRUE,  TRUE,  TRUE,
            FALSE, TRUE,  FALSE, TRUE),
  stringsAsFactors = FALSE
)
N_PARAMS <- nrow(param_bounds)

# =============================================================================
# 4.  PARAMETER APPLICATION  (one function, one HRU at a time)
# =============================================================================
apply_params_to_hru <- function(h, p){

  is_channel <- isTRUE(!is.na(h$class$endNode))

  if(!is_channel){

    # Soil-group-specific subsurface parameters
    grp <- h$class$soil
    if(is.null(grp) || is.na(grp)) grp <- 1L
    grp <- as.integer(grp)

    h$sz$parameters["t_0"] <- p[paste0("t_0_", grp)]
    h$sz$parameters["m"]   <- p[paste0("m_",   grp)]
    h$uz$parameters["t_d"] <- p[paste0("t_d_", grp)]

    # Shared hillslope params
    h$rz$parameters["s_rzmax"] <- p["s_rzmax"]
    if(isTRUE(h$sf$type == "cnst"))
      h$sf$parameters["c_sf"]  <- p["c_sf"]

  } else {

    # Channel: fixed low subsurface, surface velocity = shared c_sf
    h$sz$parameters["t_0"]     <- 0.001
    h$sz$parameters["m"]       <- 0.001
    h$rz$parameters["s_rzmax"] <- 0.001
    if(isTRUE(h$sf$type == "cnst")){
      h$sf$parameters["c_sf"]  <- p["c_sf"]
      h$sf$parameters["t_raf"] <- 999.9
    }
  }

  # Snow — shared across all HRU types
  if(!is.null(h$snow) && !is.null(h$snow$parameters)){
    h$snow$parameters["T_T"] <- p["T_T"]
    h$snow$parameters["C_f"] <- p["C_f"]
  }

  return(h)
}

# Verification helper — prints t_0/m/t_d for one HRU of each soil group
verify_params <- function(p){
  mdl_v <- readRDS(MDL_PATH)
  for(grp in 1:3){
    idx <- which(sapply(mdl_v$hru, function(h){
      isTRUE(as.integer(h$class$soil) == grp) &&
        isTRUE(is.na(h$class$endNode))
    }))[1]
    if(is.na(idx)){ cat("Soil group", grp, ": no HRU found\n"); next }
    hm <- apply_params_to_hru(mdl_v$hru[[idx]], p)
    cat(sprintf("Group %d (HRU %d): t_0=%.4f  m=%.5f  t_d=%.0f\n",
                grp, mdl_v$hru[[idx]]$id,
                hm$sz$parameters["t_0"],
                hm$sz$parameters["m"],
                hm$uz$parameters["t_d"]))
  }
}

# =============================================================================
# 5.  MODEL RUNNER
# =============================================================================
run_model <- function(p,
                      mdl_path   = MDL_PATH,
                      obs        = obs_sim,
                      outlet_id  = OUTLET_ID,
                      pa_hru_ids = sort(unique(pa_obs$hru_id)),
                      sub_step   = 3600L){

  # Reload fresh from disk every call — eliminates all state corruption risk
  mdl <- readRDS(mdl_path)

  # Zero uninitialised snow states
  for(ii in seq_along(mdl$hru)){
    sv <- mdl$hru[[ii]]$states
    if("s_snow"   %in% names(sv)) mdl$hru[[ii]]$states["s_snow"]   <- 0.0
    if("s_liquid" %in% names(sv)) mdl$hru[[ii]]$states["s_liquid"] <- 0.0
  }

  # Apply parameters — lapply returns a new list, never mutates mdl
  hru <- lapply(mdl$hru, apply_params_to_hru, p = p)

  # Initialisation
  total_area <- sum(sapply(hru, function(h) h$properties["area"]))
  q_init     <- as.numeric(na.omit(obs$Discharge_m3_s))[1]
  if(!is.finite(q_init) || q_init <= 0) q_init <- 1e-5

  for(ii in seq_along(hru)){
    is_ch  <- isTRUE(!is.na(hru[[ii]]$class$endNode))
    t0_v   <- hru[[ii]]$sz$parameters["t_0"]
    m_v    <- hru[[ii]]$sz$parameters["m"]
    grad_v <- as.numeric(hru[[ii]]$properties["gradient"])
    if(!is.finite(grad_v) || grad_v <= 0) grad_v <- 1e-4
    max_q  <- t0_v * m_v * grad_v   # max lateral drainage for this HRU

    hru[[ii]]$initialisation["s_rz_0"]    <- 0.75
    hru[[ii]]$initialisation["r_uz_sz_0"] <- min(q_init / total_area,
                                                   max_q * 0.5)
  }

  # Simulate
  out_rows <- build_out_rows(pa_hru_ids, outlet_id)
  obs_in   <- strip_obs(obs)

  result <- tryCatch({
    cm <- dynatop$new(hru, map = mdl$map)
    cm$add_data(obs_in)
    cm$initialise()
    cm$sim(out_rows, sub_step = sub_step)$get_output()
  }, error = function(e) NULL)

  return(result)
}

# =============================================================================
# 6.  OBJECTIVE FUNCTIONS
# =============================================================================
kge_fun <- function(sim, obs){
  b     <- merge(sim, obs, join = "inner")
  s     <- as.numeric(b[,1]); o <- as.numeric(b[,2])
  valid <- is.finite(s) & is.finite(o) & abs(s) < 1e3
  if(sum(valid) < 10) return(-Inf)
  s <- s[valid]; o <- o[valid]
  r <- cor(s, o); if(!is.finite(r)) return(-Inf)
  1 - sqrt((r-1)^2 + (sd(s)/sd(o)-1)^2 + (mean(s)/mean(o)-1)^2)
}

nse_fun <- function(sim, obs){
  b     <- merge(sim, obs, join = "inner")
  s     <- as.numeric(b[,1]); o <- as.numeric(b[,2])
  valid <- is.finite(s) & is.finite(o) & abs(s) < 1e3
  if(sum(valid) < 10) return(-Inf)
  s <- s[valid]; o <- o[valid]
  1 - sum((s-o)^2) / sum((o - mean(o))^2)
}

pa_metrics <- function(sim_out, pa_obs, threshold = FLOW_THRESHOLD){
  TP <- FP <- FN <- TN <- 0L
  for(hid in unique(pa_obs$hru_id)){
    col <- paste0("q_sf_hru", hid)
    if(!col %in% names(sim_out)) next
    sub <- pa_obs[pa_obs$hru_id == hid, ]
    if(nrow(sub) == 0) next
    sx  <- xts(sub$observed, order.by = as.POSIXct(sub$datetime, tz="UTC"))
    b   <- merge(sim_out[, col], sx, join = "inner")
    if(nrow(b) == 0) next
    sb  <- as.integer(as.numeric(b[,1]) >= threshold)
    ob  <- as.integer(as.numeric(b[,2]))
    ok  <- !is.na(sb) & !is.na(ob)
    sb  <- sb[ok]; ob <- ob[ok]
    TP  <- TP + sum(sb==1L & ob==1L)
    FP  <- FP + sum(sb==1L & ob==0L)
    FN  <- FN + sum(sb==0L & ob==1L)
    TN  <- TN + sum(sb==0L & ob==0L)
  }
  n <- TP + FP + FN + TN
  list(CSI      = if((TP+FP+FN)>0) TP/(TP+FP+FN) else NA_real_,
       accuracy = if(n>0) (TP+TN)/n             else NA_real_,
       TP=TP, FP=FP, FN=FN, TN=TN)
}

# =============================================================================
# 7.  LIKELIHOOD
# =============================================================================
W_KGE      <- 0.30
W_NSE      <- 0.10
W_CSI      <- 0.30
W_ACC      <- 0.30
KGE_THRESH <- -0.5    # permissive — tighten once behavioural sets appear

glue_L <- function(kge_v, nse_v, csi_v, acc_v, thresh = KGE_THRESH){
  if(!is.finite(kge_v) || kge_v < thresh) return(0)
  ks  <- (kge_v - thresh) / (1 - thresh)
  ns  <- max(0, nse_v)
  cs  <- if(is.finite(csi_v)) csi_v else 0
  ac  <- if(is.finite(acc_v)) acc_v else 0
  (ks^W_KGE) * (ns^W_NSE) * (cs^W_CSI) * (ac^W_ACC)
}

# =============================================================================
# 8.  LHS SAMPLER
# =============================================================================
generate_lhs <- function(n, pb){
  d <- randomLHS(n, nrow(pb))
  M <- matrix(NA_real_, n, nrow(pb)); colnames(M) <- pb$name
  for(i in seq_len(nrow(pb))){
    lo <- pb$lower[i]; up <- pb$upper[i]
    if(isTRUE(pb$log[i])){
      M[,i] <- 10^(d[,i]*(log10(up)-log10(lo)) + log10(lo))
    } else {
      M[,i] <- d[,i]*(up-lo) + lo
    }
  }
  M
}

# =============================================================================
# 9.  SINGLE-SAMPLE EVALUATOR  (exported to parallel workers)
# =============================================================================
evaluate_sample <- function(param_vec){
  p <- setNames(param_vec, param_bounds$name)

  # Hard guards
  if(any(p[c("t_d_1","t_d_2","t_d_3")] < 3600))
    return(list(L=0,kge=NA,nse=NA,csi=NA,accuracy=NA,bias=NA))
  if(p["s_rzmax"]<=0 || p["c_sf"]<=0)
    return(list(L=0,kge=NA,nse=NA,csi=NA,accuracy=NA,bias=NA))

  sim <- run_model(p)
  if(is.null(sim))
    return(list(L=0,kge=NA,nse=NA,csi=NA,accuracy=NA,bias=NA))

  sv <- as.numeric(sim)
  if(any(!is.finite(sv)) || any(abs(sv) > 1e3))
    return(list(L=0,kge=NA,nse=NA,csi=NA,accuracy=NA,bias=NA))

  sim_q  <- sim[,"q_out"]
  kge_v  <- kge_fun(sim_q, obs_q)
  nse_v  <- nse_fun(sim_q, obs_q)

  b      <- merge(sim_q, obs_q, join="inner")
  bias_v <- mean(as.numeric(b[,1]),na.rm=TRUE) /
    mean(as.numeric(b[,2]),na.rm=TRUE)
  if(!is.finite(bias_v) || bias_v > 100 || bias_v < 0.001)
    return(list(L=0,kge=NA,nse=NA,csi=NA,accuracy=NA,bias=bias_v))

  pa     <- pa_metrics(sim, pa_obs)
  csi_v  <- if(!is.null(pa) && is.finite(pa$CSI))      pa$CSI      else 0
  acc_v  <- if(!is.null(pa) && is.finite(pa$accuracy)) pa$accuracy else 0

  L <- glue_L(kge_v, nse_v, csi_v, acc_v)
  list(L=L, kge=kge_v, nse=nse_v, csi=csi_v, accuracy=acc_v, bias=bias_v)
}

# =============================================================================
# 10.  GLUE RUNNER
# =============================================================================
run_glue <- function(n_samples = 500,
                     n_cores   = 1L,
                     seed      = 42L,
                     pb        = param_bounds,
                     tag       = "run"){

  set.seed(seed)
  pm <- generate_lhs(n_samples, pb)
  cat(sprintf("Stage [%s]: %d samples, %d cores\n", tag, n_samples, n_cores))

  exports <- c("param_bounds","N_PARAMS",
               "MDL_PATH","obs_sim","obs_q","pa_obs",
               "OUTLET_ID","FLOW_THRESHOLD",
               "W_KGE","W_NSE","W_CSI","W_ACC","KGE_THRESH",
               "run_model","apply_params_to_hru",
               "build_out_rows","strip_obs",
               "kge_fun","nse_fun","pa_metrics","glue_L",
               "evaluate_sample")

  if(n_cores > 1L){
    cl <- makeCluster(n_cores)
    on.exit(stopCluster(cl), add=TRUE)
    clusterExport(cl, exports, envir=.GlobalEnv)
    clusterEvalQ(cl, { library(dynatop); library(xts) })
    sc <- parApply(cl, pm, 1, evaluate_sample)
  } else {
    sc <- apply(pm, 1, evaluate_sample)
  }

  res <- data.frame(pm,
                    L        = sapply(sc, `[[`, "L"),
                    kge      = sapply(sc, `[[`, "kge"),
                    nse      = sapply(sc, `[[`, "nse"),
                    csi      = sapply(sc, `[[`, "csi"),
                    accuracy = sapply(sc, `[[`, "accuracy"),
                    bias     = sapply(sc, `[[`, "bias"))

  nb <- sum(res$L > 0, na.rm=TRUE)
  cat(sprintf("  Behavioural: %d/%d (%.1f%%)  best KGE=%.3f  best NSE=%.3f\n",
              nb, n_samples, 100*nb/n_samples,
              max(res$kge, na.rm=TRUE), max(res$nse, na.rm=TRUE)))

  res$L_norm <- if(sum(res$L,na.rm=TRUE)>0) res$L/sum(res$L,na.rm=TRUE) else NA
  res <- res[order(-res$L), ]
  saveRDS(res, file.path(OUT_DIR, paste0("glue_",tag,".rds")))
  res
}

# =============================================================================
# 11.  TIGHTEN BOUNDS
# =============================================================================
tighten_bounds <- function(pb, behav, prob=c(0.05,0.95), pad=1.2){
  pb2 <- pb
  for(i in seq_len(nrow(pb))){
    nm <- pb$name[i]; vals <- behav[[nm]]
    if(length(vals) < 5) next
    q  <- quantile(vals, probs=prob, na.rm=TRUE)
    if(isTRUE(pb$log[i])){
      lo <- max(q[1]/pad, 1e-7); up <- q[2]*pad
    } else {
      w  <- (q[2]-q[1])*(pad-1)/2; lo <- q[1]-w; up <- q[2]+w
    }
    pb2$lower[i] <- max(pb$lower[i], lo)
    pb2$upper[i] <- min(pb$upper[i], up)
  }
  pb2
}

# =============================================================================
# 12.  DOTTY PLOTS
# =============================================================================
plot_dotty <- function(results, pb, metric="L"){
  behav <- results[results$L > 0 & !is.na(results$L), ]
  nr    <- ceiling(sqrt(nrow(pb))); nc <- ceiling(nrow(pb)/nr)
  par(mfrow=c(nr,nc))
  for(nm in pb$name){
    ymax <- if(nrow(behav)>0) max(behav[[metric]],na.rm=TRUE) else 1
    xlim <- c(pb$lower[pb$name==nm], pb$upper[pb$name==nm])
    plot(results[[nm]], results[[metric]],
         xlab=nm, ylab=metric, pch=16, cex=0.4,
         col="grey70", ylim=c(0,ymax), xlim=xlim, main=nm)
    if(nrow(behav)>0)
      points(behav[[nm]], behav[[metric]], pch=16, cex=0.7, col="firebrick")
    abline(h=0, lty=2, col="grey50")
  }
  par(mfrow=c(1,1))
}

# =============================================================================
# 13.  EXECUTE
# =============================================================================

# ── Step 0: verify parameters actually change ─────────────────────────────────
cat("=== Parameter verification (low vs high bounds) ===\n")
p_lo <- setNames(param_bounds$lower, param_bounds$name)
p_hi <- setNames(param_bounds$upper, param_bounds$name)
p_lo[c("t_d_1","t_d_3")] <- 7200; p_lo["t_d_2"] <- 90000
cat("--- Low params ---\n");  verify_params(p_lo)
cat("--- High params ---\n"); verify_params(p_hi)

# ── Step 1: single test run ───────────────────────────────────────────────────
cat("\n=== Test run (geometric midpoint) ===\n")
p_mid <- p_lo
p_mid[param_bounds$log]  <- sqrt(param_bounds$lower[param_bounds$log] *
                                   param_bounds$upper[param_bounds$log])
p_mid[!param_bounds$log] <- (param_bounds$lower[!param_bounds$log] +
                                param_bounds$upper[!param_bounds$log]) / 2

t0       <- Sys.time()
test_sim <- run_model(p_mid)
cat(sprintf("Runtime: %.1fs\n", as.numeric(Sys.time()-t0, units="secs")))

if(!is.null(test_sim)){
  cat("KGE:", round(kge_fun(test_sim[,"q_out"], obs_q), 3), "\n")
  cat("NSE:", round(nse_fun(test_sim[,"q_out"], obs_q), 3), "\n")
  pa_t <- pa_metrics(test_sim, pa_obs)
  cat(sprintf("CSI=%.3f  Acc=%.3f  TP=%d FP=%d FN=%d TN=%d\n",
              pa_t$CSI, pa_t$accuracy, pa_t$TP, pa_t$FP, pa_t$FN, pa_t$TN))
  plot(merge(obs_q, test_sim[,"q_out"]),
       main="Test run", col=c("black","firebrick"), legend.loc="topright")
} else {
  stop("Test run returned NULL — fix run_model before proceeding")
}

# ── Step 2: broad Stage 1 screen ──────────────────────────────────────────────
Sys.time()
g1 <- run_glue(n_samples = 2000, n_cores = 6L, seed = 42L,
               pb = param_bounds, tag = "stage1")
Sys.time()
saveRDS(g1, file.path(OUT_DIR, "unfocused_2000_seed42.rds"))

# LOOP
for(z in 1:10){
g1 <- run_glue(n_samples = 1000, n_cores = 6L, seed = 50L+z,
               pb = param_bounds, tag = "stage1")
#Sys.time()
saveRDS(g1,
        paste0(OUT_DIR, "/glue_screen1000_seed",50+z,".rds"))
print(paste0("run ",z," out of 48"))

}

files <- list.files(OUT_DIR)
for(i in 1:length(files)){
  file <- files[i]
  
  raw_glue <- readRDS(paste0(OUT_DIR, "/",file))
  screen_glue <- raw_glue[raw_glue$L >= 0 & !is.na(raw_glue$L), ]
  
  if(i == 1) alldat <- screen_glue
  if(i > 1) alldat <- rbind(alldat, screen_glue)
}
screen_behav <- alldat %>% 
  filter(kge >= -0.4)
screen_behav[1:10,]
screen_behav %>% 
  tibble() %>% 
arrange(desc(kge))
g1 <- screen_behav
behav1 <- g1[g1$L > 0 & !is.na(g1$L), ]

cat("\nKGE quantiles from Stage 1:\n")
print(quantile(g1$kge, c(0,.1,.25,.5,.75,.9,1), na.rm=TRUE))

# ── Step 3: tighten and refine ────────────────────────────────────────────────
# Use behavioural runs if available, otherwise top 10% by KGE
#behav1 <- g1[g1$L > 0 & !is.na(g1$L), ]
focus  <- if(nrow(behav1) >= 10) behav1 else g1[order(-g1$kge),][1:max(10,nrow(g1)%/%10),]
pb2    <- tighten_bounds(param_bounds, focus, prob=c(0.05,0.95), pad=1.3)

cat("\nFocused bounds:\n")
print(pb2[, c("name","lower","upper")])

g2 <- run_glue(n_samples = 3000, n_cores = 7L, seed = 99L,
               pb = pb2, tag = "stage2")
saveRDS(g1, file.path(OUT_DIR, "stage2_3000_seed99.rds"))

for(z in 1:15){
g1 <- run_glue(n_samples = 1000, n_cores = 7L, seed = 90L+z,
               pb = pb2, tag = "stage2")
#Sys.time()
saveRDS(g1,
        paste0("./7th_attempt/glue_results_focused/glue_screen1000_seed",90+z,".rds"))
print(paste0("run ",z," out of 15"))

}

files <- list.files("./7th_attempt/glue_results_focused")
for(i in 1:length(files)){
  file <- files[i]
  
  raw_glue <- readRDS(paste0("./7th_attempt/glue_results_focused/",file))
  screen_glue <- raw_glue[raw_glue$L >= 0.5 & !is.na(raw_glue$L), ]
  
  if(i == 1) alldat <- screen_glue
  if(i > 1) alldat <- rbind(alldat, screen_glue)
}

good_ones <- alldat
# ── Step 4: combine and evaluate best ────────────────────────────────────────
good_ones[order(-good_ones$kge), ]

all_res <- good_ones[order(-good_ones$L), ]
all_res$L_norm <- all_res$L / sum(all_res$L, na.rm=TRUE)
saveRDS(all_res, file.path(OUT_DIR, "glue_all.rds"))
all_res[1,]
cat("\n=== Best parameter set ===\n")
best_p <- setNames(as.numeric(all_res[1, param_bounds$name]), param_bounds$name)
print(round(best_p, 6))
cat(sprintf("L=%.4f  KGE=%.3f  NSE=%.3f  CSI=%.3f  Acc=%.3f  Bias=%.3f\n",
            all_res$L[1], all_res$kge[1], all_res$nse[1],
            all_res$csi[1], all_res$accuracy[1], all_res$bias[1]))

# Best Lnorm
best_sim <- run_model(best_p)
if(!is.null(best_sim)){
  plot(merge(obs_q, best_sim[,"q_out"]),
       main="Best run: observed vs simulated [m3/s]",
       col=c("black","firebrick"), legend.loc="topright")
}
# Best kge
best_kge <- good_ones[order(-good_ones$kge), ][1,]
best_p <- setNames(as.numeric(best_kge[1, param_bounds$name]), param_bounds$name)

best_sim <- run_model(best_p)
#if(!is.null(best_sim)){
  plot(merge(best_sim[,"q_out"], obs_q),
       main="Best run: observed vs simulated [m3/s]",
       col=c("firebrick", "black"), legend.loc="topright")
# ── Step 5: dotty plots ───────────────────────────────────────────────────────
plot_dotty(all_res, param_bounds, metric="L")
```


```{r}
test_parms <- data.frame(
  name = param_bounds$name,
  value = c(0.135, # t_0
            0.00698, # m
            8*60*60, # t_d
            0.1, # s_rzmax
            0.5425, #c_sf
            0.05, #n_ch
            1, # v_ch
            0, # T_T
            2.5, # C_f
            0.1 # C_wh
            )
)
test_params <- setNames(
  test_parms$value,
  test_parms$name
)

test_sim <- run_model(test_params, mdl, obs_sim,
                      outlet_id  = OUTLET_ID,
                      pa_hru_ids = integer(0),   # skip PA to keep it clean
                      include_sz = FALSE)

plot(merge(test_sim, obs_sim[,1]))
kge(test_sim[,1], obs_sim)
```



# PLOT
```{r}
# =============================================================================
# Figure 1: Spatial map of per-sensor accuracy
# =============================================================================

library(ggplot2)
library(terra)
library(tidyverse)
library(sf)

# ── Compute per-sensor accuracy from best_sim ─────────────────────────────────
compute_sensor_accuracy <- function(best_sim, pa_obs, 
                                    threshold = FLOW_THRESHOLD){
  do.call(rbind, lapply(unique(pa_obs$hru_id), function(hid){
    col_name <- paste0("q_sf_hru", hid)
    if(!col_name %in% names(best_sim)) return(NULL)
    
    sub <- pa_obs[pa_obs$hru_id == hid, ]
    if(nrow(sub) == 0) return(NULL)
    
    sub_xts <- xts(sub$observed,
                   order.by = as.POSIXct(sub$datetime, tz = "UTC"))
    both    <- merge(best_sim[, col_name], sub_xts, join = "inner")
    if(nrow(both) == 0) return(NULL)
    
    sim_bin <- as.integer(as.numeric(both[, 1]) >= threshold)
    obs_bin <- as.integer(as.numeric(both[, 2]))
    valid   <- !is.na(sim_bin) & !is.na(obs_bin)
    sim_bin <- sim_bin[valid]; obs_bin <- obs_bin[valid]
    
    TP <- sum(sim_bin == 1L & obs_bin == 1L)
    FP <- sum(sim_bin == 1L & obs_bin == 0L)
    FN <- sum(sim_bin == 0L & obs_bin == 1L)
    TN <- sum(sim_bin == 0L & obs_bin == 0L)
    n  <- TP + FP + FN + TN
    
    data.frame(
      hru_id   = hid,
      accuracy = if(n > 0) (TP + TN) / n else NA_real_,
      CSI      = if((TP+FP+FN) > 0) TP / (TP+FP+FN) else NA_real_,
      POD      = if((TP+FN) > 0) TP / (TP+FN) else NA_real_,
      FAR      = if((TP+FP) > 0) FP / (TP+FP) else NA_real_,
      TP=TP, FP=FP, FN=FN, TN=TN, n=n,
      pct_flowing = mean(obs_bin, na.rm=TRUE)
    )
  }))
}

sensor_acc <- compute_sensor_accuracy(best_sim, pa_obs)
locs
locs_df <- as.data.frame(stic_locs, geom = "XY") %>%
  rename(sensor_id = number, lon = x, lat = y) %>%
  group_by(sensor_id) %>%
  #slice(2) %>%          # keep first occurrence — or use summarise(lon=mean(lon), lat=mean(lat))
  ungroup() %>% 
  distinct()

#spatial_acc <- 
  sensor_hru_key %>%
  left_join(sensor_acc, by = "hru_id") %>%
  left_join(locs_df, by = "sensor_id") %>%
  filter(!is.na(accuracy), !is.na(lon)) #%>%
  distinct(sensor_id, .keep_all = TRUE)  # belt-and-suspenders dedup

# Join sensor accuracy to spatial locations via sensor_hru_map
# sensor_hru_map has columns: sensor_id, hru_id, hru_name, seg_length, snap_dist
spatial_acc <- 
sensor_hru_key %>%
  left_join(sensor_acc, by = "hru_id") %>%
  left_join(locs_df) %>% 
  filter(!is.na(accuracy), !is.na(lon))

# Convert to sf for plotting
spatial_acc_sf <- st_as_sf(spatial_acc,
                            coords = c("lon", "lat"),
                            crs = "EPSG:26919")

# Watershed outline and channel for basemap
watershed_sf <- st_as_sf(w3_outline)     # adjust to your outline object
channel_sf   <- st_as_sf(vect_final)

# ── Plot ──────────────────────────────────────────────────────────────────────
fig1 <- ggplot() +
  geom_sf(data = watershed_sf, fill = "grey96", color = "grey60",
          linewidth = 0.4) +
  geom_sf(data = channel_sf, color = "steelblue", linewidth = 0.5,
          alpha = 0.7) +
  geom_sf(data = spatial_acc_sf,
          aes(fill = accuracy),
          shape = 21, color = "white", stroke = 0.6, size = 6) +
  scale_fill_distiller(
    palette  = "RdYlGn",
    direction = 1,
    limits   = c(0, 1),
    name     = "Accuracy",
    labels   = scales::percent_format(accuracy = 1)
  ) +
  # scale_size_continuous(
  #   name   = "Observations",
  #   range  = c(3, 9),
  #   breaks = c(50, 200, 500, 1000)
  # ) +
  # Label each point with sensor id and accuracy %
  ggrepel::geom_label_repel(
    data = spatial_acc_sf,
    aes(label = paste0(sensor_id, "\n",
                        scales::percent(accuracy, accuracy = 1)),
        geometry = geometry),
    stat         = "sf_coordinates",
    size         = 2.8,
    label.size   = 0.2,
    label.padding = unit(0.15, "lines"),
    box.padding  = unit(0.4, "lines"),
    min.segment.length = 0.3,
    fill         = "white",
    color        = "grey20"
  ) +
  coord_sf() +
  labs(
    title    = "Flow presence/absence accuracy by sensor location",
    subtitle = paste0("Calibration period  ",
                       format(cal_start, "%Y-%m-%d"), " — ",
                       format(cal_end,   "%Y-%m-%d")),
    caption  = paste0("Threshold: ", FLOW_THRESHOLD, " m³/s  |  ",
                       "n = observations per site")
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title    = element_text(size = 13, face = "bold"),
    plot.subtitle = element_text(color = "grey50"),
    legend.position = "right",
    panel.grid    = element_line(color = "grey92", linewidth = 0.3)
  )

print(fig1)
ggsave("fig1_spatial_accuracy.png", fig1, scale = 2,
       dpi = 300, bg = "white")

# =============================================================================
# Figure 2: Accuracy through time, aligned with hydrograph
# =============================================================================
# Computes a rolling accuracy for each sensor over a sliding window,
# then stacks them in a panel below the hydrograph.


# Hydrograph data
hydro_df <- data.frame(
  datetime = index(obs_q),
  q_obs    = as.numeric(obs_q),
  q_sim    = as.numeric(best_sim[index(obs_q), "q_out"])
) %>%
  filter(datetime >= min(rolling_acc$datetime, na.rm=TRUE),
         datetime <= max(rolling_acc$datetime, na.rm=TRUE))

# ── Panel 1: hydrograph ───────────────────────────────────────────────────────
p_hydro <- ggplot(hydro_df, aes(x = datetime)) +
  geom_line(aes(y = q_obs, color = "Observed"), linewidth = 0.5) +
  geom_line(aes(y = q_sim, color = "Simulated"), linewidth = 0.5,
            alpha = 0.8) +
  scale_color_manual(values = c("Observed" = "black",
                                 "Simulated" = "firebrick"),
                     name = NULL) +
  scale_y_continuous(labels = scales::label_number(suffix = " m³/s",
                                                    accuracy = 0.001)) +
  labs(y = "Discharge", x = NULL) +
  theme_minimal(base_size = 10) +
  theme(
    legend.position  = c(0.85, 0.85),
    legend.key.size  = unit(0.5, "cm"),
    panel.grid.minor = element_blank(),
    axis.text.x      = element_blank(),
    axis.ticks.x     = element_blank(),
    plot.margin      = margin(t=8, b=0, l=8, r=8)
  )+xlim(c(window1_start, window1_end))

# ── Panel 2: rolling accuracy per sensor, faceted ────────────────────────────

compute_rolling_accuracy <- function(best_sim, pa_obs,
                                     threshold    = FLOW_THRESHOLD,
                                     bin_days     = 14,      # width of each accuracy bin
                                     join_tol_min = 35){     # max minutes to match obs to sim
  
  # ── 1. Convert sim to data frame once — avoid repeated xts indexing ─────────
  sim_df <- data.frame(
    datetime = as.POSIXct(index(best_sim), tz = "UTC"),
    as.data.frame(coredata(best_sim))
  )
  
  # ── 2. Tolerance join: snap each pa_obs row to nearest sim timestep ──────────
  # Uses findInterval — O(n log n), no loops
  tol_secs <- join_tol_min * 60
  
  pa_joined <- pa_obs %>%
    mutate(datetime = as.POSIXct(datetime, tz = "UTC")) %>%
    mutate(
      # Index of nearest sim timestep
      idx = findInterval(as.numeric(datetime),
                         as.numeric(sim_df$datetime)),
      idx = pmax(1L, pmin(idx, nrow(sim_df)))
    ) %>%
    mutate(
      sim_time = sim_df$datetime[idx],
      gap_secs = abs(as.numeric(datetime) - as.numeric(sim_time))
    ) %>%
    filter(gap_secs <= tol_secs) %>%   # drop obs with no nearby sim timestep
    mutate(sim_idx = idx)
  
  cat(sprintf("PA obs matched: %d / %d (tol = %d min)\n",
              nrow(pa_joined), nrow(pa_obs), join_tol_min))
  
  # ── 3. For each HRU, pull simulated values and compute binned accuracy ───────
  all_results <- do.call(rbind, lapply(unique(pa_joined$hru_id), function(hid){
    
    col_name <- paste0("q_sf_hru", hid)
    if(!col_name %in% names(sim_df)) return(NULL)
    
    sub <- pa_joined %>%
      filter(hru_id == hid) %>%
      mutate(
        sim_q   = sim_df[[col_name]][sim_idx],
        sim_bin = as.integer(sim_q >= threshold),
        obs_bin = as.integer(observed),
        # Assign each obs to a calendar bin
        bin_start = as.POSIXct(
          floor(as.numeric(datetime) / (bin_days * 86400)) * (bin_days * 86400),
          origin = "1970-01-01", tz = "UTC"
        )
      ) %>%
      filter(!is.na(sim_bin), !is.na(obs_bin))
    
    if(nrow(sub) == 0) return(NULL)
    
    # ── 4. Summarise per bin — fully vectorized, no inner loop ─────────────────
    sub %>%
      group_by(hru_id, bin_start) %>%
      summarise(
        TP       = sum(sim_bin == 1L & obs_bin == 1L),
        FP       = sum(sim_bin == 1L & obs_bin == 0L),
        FN       = sum(sim_bin == 0L & obs_bin == 1L),
        TN       = sum(sim_bin == 0L & obs_bin == 0L),
        n_obs    = n(),
        .groups  = "drop"
      ) %>%
      mutate(
        accuracy = (TP + TN) / (TP + FP + FN + TN),
        CSI      = ifelse((TP+FP+FN) > 0, TP / (TP+FP+FN), NA_real_),
        POD      = ifelse((TP+FN)    > 0, TP / (TP+FN),    NA_real_),
        datetime = bin_start + (bin_days * 86400 / 2)  # bin midpoint for plotting
      ) %>%
      filter(n_obs >= 3)   # skip bins with too few observations to be meaningful
  }))
  
  all_results
}

# ── Usage ─────────────────────────────────────────────────────────────────────
rolling_acc <- compute_rolling_accuracy(
  best_sim     = best_sim,
  pa_obs       = pa_obs,
  threshold    = FLOW_THRESHOLD,
  bin_days     = 14,       # 2-week bins — reduce to 7 for finer resolution
  join_tol_min = 20        # 20-min tolerance handles both your 10-min and 30-min series
)
# Sort sensors by upstream-to-downstream order (hru_id ascending)
rolling_acc <- rolling_acc %>% 
  left_join(sensor_hru_map %>% select(sensor_id, hru_id),
            by = "hru_id") %>%
  mutate(label = paste0("Site ", sensor_id, " (HRU ", hru_id, ")")) %>%  # create first
  mutate(label = fct_reorder(label, hru_id))                              # then reorder


# Tag each row with which period it belongs to
window1_start <- ymd("2023-07-01", tz = "UTC")
window1_end   <- ymd("2023-11-30", tz = "UTC")
window2_start <- as.POSIXct("2024-05-01", tz = "UTC")
window2_end   <- as.POSIXct("2024-07-31", tz = "UTC")
tag_period <- function(df){
  df %>% mutate(
    period = case_when(
      datetime >= window1_start & datetime <= window1_end ~ "Summer–Fall 2023",
      datetime >= window2_start & datetime <= window2_end ~ "Summer 2024",
      TRUE ~ NA_character_
    )
  ) %>% filter(!is.na(period))
}

rolling_acc <- tag_period(rolling_acc)
hydro_df       <- tag_period(hydro_df)

# Then add to both p_hydro and p_acc:
# + facet_wrap(~ period, scales = "free_x", nrow = 1)
# + scale_x_datetime(date_labels = "%b", date_breaks = "1 month")

p_acc <- ggplot(rolling_acc, aes(x = datetime, y = accuracy)) +
  # Shaded "good" zone
  annotate("rect", xmin=-Inf, xmax=Inf, ymin=0.7, ymax=1,
           fill = "green", alpha = 0.04) +
  geom_hline(yintercept = 0.5, linetype = "dashed",
             color = "grey70", linewidth = 0.4) +
  geom_line(aes(color = accuracy, group = 1), linewidth = 0.6,
            na.rm = TRUE) +
  geom_point(aes(color = accuracy), size = 0.8, na.rm = TRUE) +
  scale_color_distiller(
    palette   = "RdYlGn",
    direction = 1,
    limits    = c(0, 1),
    name      = "Accuracy",
    labels    = scales::percent_format(accuracy = 1),
    guide     = guide_colorbar(barwidth = 0.5, barheight = 5)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = scales::percent_format(accuracy = 1),
    breaks = c(0, 0.5, 1)
  ) +
  scale_x_datetime(date_labels = "%b %Y", date_breaks = "2 months",
                   limits = c(window1_start, window1_end)) +
  facet_wrap(~ label, ncol = 1, strip.position = "right") +
  labs(y = paste0("Rolling accuracy\n(", 14, "-day window)"), x = NULL) +
  theme_minimal(base_size = 9) +
  theme(
    strip.text       = element_text(size = 8, color = "grey40"),
    panel.grid.minor = element_blank(),
    panel.spacing    = unit(0.15, "lines"),
    legend.position  = "right",
    plot.margin      = margin(t=0, b=8, l=8, r=8)
  )

# ── Combine with patchwork ────────────────────────────────────────────────────
library(patchwork)

n_sensors  <- length(unique(rolling_acc$hru_id))
hydro_h    <- 2      # inches for hydrograph panel
sensor_h   <- 0.9   # inches per sensor panel

fig2 <- p_hydro / p_acc +
  plot_layout(
    heights = c(hydro_h, sensor_h * n_sensors)
  ) +
  plot_annotation(
    title   = "Rolling flow presence/absence accuracy through time",
    subtitle = paste0("14-day sliding window  |  threshold = ",
                       FLOW_THRESHOLD, " m³/s"),
    theme = theme(
      plot.title    = element_text(size = 13, face = "bold"),
      plot.subtitle = element_text(size = 10, color = "grey50")
    )
  )
fig2
print(fig2)
ggsave("fig2_temporal_accuracy.png", fig2,
       width  = 10,
       height = hydro_h + sensor_h * n_sensors + 1.5,
       dpi    = 300, bg = "white")
```

```{r}
compute_rolling_accuracy <- function(best_sim, pa_obs,
                                     threshold    = FLOW_THRESHOLD,
                                     bin_days     = 14,      # width of each accuracy bin
                                     join_tol_min = 20){     # max minutes to match obs to sim
  
  # ── 1. Convert sim to data frame once — avoid repeated xts indexing ─────────
  sim_df <- data.frame(
    datetime = as.POSIXct(index(best_sim), tz = "UTC"),
    as.data.frame(coredata(best_sim))
  )
  
  # ── 2. Tolerance join: snap each pa_obs row to nearest sim timestep ──────────
  # Uses findInterval — O(n log n), no loops
  tol_secs <- join_tol_min * 60
  
  pa_joined <- pa_obs %>%
    mutate(datetime = as.POSIXct(datetime, tz = "UTC")) %>%
    mutate(
      # Index of nearest sim timestep
      idx = findInterval(as.numeric(datetime),
                         as.numeric(sim_df$datetime)),
      idx = pmax(1L, pmin(idx, nrow(sim_df)))
    ) %>%
    mutate(
      sim_time = sim_df$datetime[idx],
      gap_secs = abs(as.numeric(datetime) - as.numeric(sim_time))
    ) %>%
    filter(gap_secs <= tol_secs) %>%   # drop obs with no nearby sim timestep
    mutate(sim_idx = idx)
  
  cat(sprintf("PA obs matched: %d / %d (tol = %d min)\n",
              nrow(pa_joined), nrow(pa_obs), join_tol_min))
  
  # ── 3. For each HRU, pull simulated values and compute binned accuracy ───────
  all_results <- do.call(rbind, lapply(unique(pa_joined$hru_id), function(hid){
    
    col_name <- paste0("q_sf_hru", hid)
    if(!col_name %in% names(sim_df)) return(NULL)
    
    sub <- pa_joined %>%
      filter(hru_id == hid) %>%
      mutate(
        sim_q   = sim_df[[col_name]][sim_idx],
        sim_bin = as.integer(sim_q >= threshold),
        obs_bin = as.integer(observed),
        # Assign each obs to a calendar bin
        bin_start = as.POSIXct(
          floor(as.numeric(datetime) / (bin_days * 86400)) * (bin_days * 86400),
          origin = "1970-01-01", tz = "UTC"
        )
      ) %>%
      filter(!is.na(sim_bin), !is.na(obs_bin))
    
    if(nrow(sub) == 0) return(NULL)
    
    # ── 4. Summarise per bin — fully vectorized, no inner loop ─────────────────
    sub %>%
      group_by(hru_id, bin_start) %>%
      summarise(
        TP       = sum(sim_bin == 1L & obs_bin == 1L),
        FP       = sum(sim_bin == 1L & obs_bin == 0L),
        FN       = sum(sim_bin == 0L & obs_bin == 1L),
        TN       = sum(sim_bin == 0L & obs_bin == 0L),
        n_obs    = n(),
        .groups  = "drop"
      ) %>%
      mutate(
        accuracy = (TP + TN) / (TP + FP + FN + TN),
        CSI      = ifelse((TP+FP+FN) > 0, TP / (TP+FP+FN), NA_real_),
        POD      = ifelse((TP+FN)    > 0, TP / (TP+FN),    NA_real_),
        datetime = bin_start + (bin_days * 86400 / 2)  # bin midpoint for plotting
      ) %>%
      filter(n_obs >= 3)   # skip bins with too few observations to be meaningful
  }))
  
  all_results
}

# ── Usage ─────────────────────────────────────────────────────────────────────
rolling_acc <- compute_rolling_accuracy(
  best_sim     = best_sim,
  pa_obs       = pa_obs,
  threshold    = FLOW_THRESHOLD,
  bin_days     = 14,       # 2-week bins — reduce to 7 for finer resolution
  join_tol_min = 20        # 20-min tolerance handles both your 10-min and 30-min series
)
```

