---
title: "HB_dynatop2"
format: html
editor_options: 
  chunk_output_type: console
---
Attempt 2, using 15 minute timestep inputs over a shorter period of time.
# SETUP
```{r}
library(pacman)

p_load(terra, dynatop, dynatopGIS, xts, whitebox, tidyverse,
       lubridate,
       sf, # working with st and sf objects
       tidyterra, # do tidy ops on terra objects
       sfnetworks # create start and end nodes for channel network
       )
setwd("./hb_data")
```

# GIS PREP
```{r GIS files}
dem_path <- file.path(".", "unprocessed",
                      "1m hydro enforced dem","hydem1mlpns.tif")
dem <- rast(dem_path) # load the dem as a raster layer
plot(dem)

# crop DEM to a rough rectangle
ybounds <- c(4870350,4871350)
xbounds <- c(281350, 282150)
crop1 <- crop(dem, ext(c(xbounds, ybounds)))
plot(crop1)
crop_dem_path <- file.path(".", "unprocessed",
                      "1m hydro enforced dem","hydem1mlpns_cropped.tif")
writeRaster(crop1, crop_dem_path, overwrite = TRUE)
####delineate watershed and keep watershed boundary

w3_breached <- "./unprocessed/whitebox_outputs/1mdem_breach.tif"
wbt_breach_depressions_least_cost(
  dem = crop_dem_path,
  output = w3_breached,
  dist = 1,
  fill = TRUE)

w3_filled <- "./unprocessed/whitebox_outputs/1mdem_fill.tif"
wbt_fill_depressions_wang_and_liu(
  dem = w3_breached,
  output = w3_filled
)
#calculate flow accumulation and direction
w3_flowacc <- "./unprocessed/whitebox_outputs/1mdem_w3_flowacc.tif"
wbt_d8_flow_accumulation(input = w3_filled,
                         output = w3_flowacc)
w3_d8pt <- "./unprocessed/whitebox_outputs/1mdem_w3_d8pt.tif"
wbt_d8_pointer(dem = w3_filled,
               output = w3_d8pt)

#delineate streams
w3_streams <- "./unprocessed/whitebox_outputs/w3_streams.tif"
wbt_extract_streams(flow_accum = w3_flowacc,
                    output = w3_streams,
                    threshold = 8000)
plot(rast(w3_streams))

# convert streams to shapefile
w3_streams_vect <- "./unprocessed/whitebox_outputs/w3_streams_vect.shp"
wbt_raster_streams_to_vector(streams = w3_streams,
                             d8_pntr = w3_d8pt,
                             output = w3_streams_vect)

# get pour point ready
pour_point_path <- file.path(".", 
                             "unprocessed",
                             "1m hydro enforced dem", "w3_pour_snap.shp")
plot(vect(pour_point_path), add = TRUE)

w3_pour_snap <- "./unprocessed/whitebox_outputs/w3_pour_snap.shp"
wbt_jenson_snap_pour_points(pour_pts = pour_point_path,
                            streams = w3_streams,
                            output = w3_pour_snap,
                            snap_dist = 10)
w3_pour_snap_read <- vect(w3_pour_snap)
plot(rast(w3_streams), 
     xlim = c(281400, 282000),
     ylim = c(4870400, 4870800))
points(w3_pour_snap_read, pch = 1)


w3_shed <- "./unprocessed/whitebox_outputs/w3_shed.tif"
wbt_watershed(d8_pntr = w3_d8pt,
              pour_pts = w3_pour_snap,
              output = w3_shed)

plot(rast(w3_shed))
#convert raster of watershed area to vector for final mapping
w3_outline <- as.polygons(rast(w3_shed), extent=FALSE)
plot(w3_outline)

plot(w3_outline)

channel <- vect(w3_streams_vect) %>% 
  crop(w3_outline)


# crop and mask DEM by watershed outline
dem_crop_mask <- rast(w3_filled) %>% 
  crop(w3_outline) %>% 
  mask(w3_outline)

#create watershed area, just use area of DEM
outline <- (dem_crop_mask * 0) + 1
# extend both rasters by 1 pixel
outline_extended <- extend(outline,1)
dem_final <- extend(dem_crop_mask,1)


writeRaster(outline_extended,file.path(".","processed","w3.tif"),overwrite=TRUE)
writeRaster(dem_final,file.path(".","processed","dem.tif"),overwrite=TRUE)
```
```{r convert channel}
plot(channel)
channel <- channel %>% 
  mutate(ID = seq(1, n(), 1))
# 2. Break multipart lines into individual components
# This converts MULTILINESTRING to LINESTRING
single_parts <- st_cast(as_sf(channel), "LINESTRING")

# 3. View the new, broken-down data
print(single_parts)
plot(single_parts)
coords <- st_coordinates(single_parts)


net <- as_sfnetwork(single_parts, directed = TRUE)

# Extract nodes
nodes <- st_as_sf(net, "nodes")

# Extract edges
edges <- st_as_sf(net, "edges")

edges$length_m <- st_length(edges)
test <- 
  edges %>% 
  select(from, to, geometry, length_m) %>% 
  mutate("Identifier" = dplyr::row_number()) %>% 
  vect() #%>%

crs(test) <- crs(dem_final)

conv_channel <- convert_channel(
    test,
    property_names = c(name = "Identifier", 
                       length = "length_m", startNode = "from",
                       endNode= "to"),
    default_width = 1,
    default_slope = 0.001
)

writeVector(conv_channel,file.path(".","processed","channel.shp"),overwrite=TRUE)

# name is different than the start and end nodes!
conv_channel %>% 
  as.data.frame() %>% 
  tibble()
```

```{r}
# Find the outlet ID right after you create your 'edges' object
edges_df <- st_drop_geometry(edges)
outlet_edge <- edges_df %>% filter(!(to %in% from))
outlet_id <- outlet_edge$ID

# Update your output definition using the correct outlet ID
output_defn <- data.frame(name = "W3",
                          id = outlet_id,
                          flux = "q_sf",
                          scale = 1)


## create a vector of the uid's of the HRUs
hid <- sapply(mdl$hru,function(h){h$id})
## create a vector of the channel names with as storaged in the class variabel
cname <- sapply(mdl$hru,function(h){h$class$name})
idx <- match(gauges$chn_identi, cname) ## index of HRU's corresponding to gauges
## create as a data.frame
gauge_output <- data.frame(name = gauges$Site.Name,
                           id = as.integer(hid[idx]),
                           flux = "q_sf",
                           scale = 1)
saveRDS(gauge_output,file.path(".","dyna","gauge_output_defn.rds")) ## save the output for later use
gauge_output
```


# TIMESERIES
```{r}
start <- "2018-05-21"
end <- "2018-10-21"

summary(flow_15min)
summary(precip)
summary(air)

#2017 - 2024
```

```{r discharge}
# discharge every 15 mins
flow_15min <- read_csv("https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/1/18/1b101627265d0b5763fa0bba93d93bf7") 

flow_xts <- 
  flow_15min %>%
  filter(DATETIME >= start,
         DATETIME <= end) %>% 
  # Create a new column that groups time into 15-minute buckets
  mutate(time_15 = floor_date(DATETIME, "15 mins")) %>%
  # Group by those buckets
  group_by(time_15) %>%
  # Average the 5-minute instantaneous rates to get the 15-minute rate
  summarize(Discharge_ls = mean(Discharge_ls, na.rm = TRUE)) %>%
  # convert to m3_s, or should it be m3/day?>
  mutate(Discharge_m3_s = Discharge_ls / 1000) %>% 
  as.xts()

```
Units of precipitation and pet are expected to be m accrued over the previous timestep.
```{r precipitation}
precip <- read_csv("https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/278/7/132ffe0622ff3f30505c774d24a6b65d")
# precip in mm accrued over timestep, needs to be m accrued over time step

precip_xts <- 
  precip %>% 
  filter(DateTime >= start,
         DateTime <= end) %>% 
  select(DateTime, precip) %>% 
  mutate(precip = precip / 1000) %>% # convert to m
  as.xts()
```

```{r pet}
# air temp already 15 min increment
air <- read_csv("https://pasta.lternet.edu/package/data/eml/knb-lter-hbr/322/7/a96172a26f381b9ec7005144d12e0820")

air_formatted <- air %>% 
  filter(DateTime >= start,
         DateTime <= end) %>% 
  select(DateTime, airTemp) 

# convert airTemp to pet with hargreaves

lat <- 45   # latitude in decimal degrees (e.g., Blacksburg VA)

# --- Hargreaves PET -----------------------------------------------------

hargreaves_pet <- function(date, temp, lat){

  J <- yday(date)                 # day of year
  phi <- lat * pi/180             # latitude radians
  
  # extraterrestrial radiation (Ra)
  dr  <- 1 + 0.033*cos(2*pi*J/365)
  delta <- 0.409*sin(2*pi*J/365 - 1.39)
  ws <- acos(-tan(phi)*tan(delta))

  Gsc <- 0.0820   # MJ m-2 min-1
  
  Ra <- (24*60/pi) * Gsc * dr *
        (ws*sin(phi)*sin(delta) +
         cos(phi)*cos(delta)*sin(ws))

  PET <- 0.0023 * Ra *
         (temp + 17.8) *
         sqrt(pmax(temp, 0))

  return(PET)   # mm/day
}

# --- Calculate PET ------------------------------------------------------

air_formatted$PET_mm_day <- hargreaves_pet(
  air_formatted$DateTime,
  air_formatted$airTemp,
  lat
)


pet_xts <- air_formatted %>% 
  select(DateTime, PET_mm_day) %>% 
  mutate(PET_m_day = PET_mm_day / (1000 * 96)) %>% # convert to m, and 15 min
  rename(pet = PET_m_day) %>% 
  select(DateTime, pet) %>% 
  as.xts()

pet_xts <- air_formatted %>% 
  select(DateTime, PET_mm_day) %>% 
  mutate(PET_m_day = PET_mm_day/1000/1000/96) %>% # convert to m, and 15 min
  rename(pet = PET_m_day) %>% 
  select(DateTime, pet) %>% 
  as.xts()

 
```

```{r combine}

obs <- merge(flow_xts,
             precip_xts,
             pet_xts,
             all=FALSE) ## merge by time stamp
#obs <- obs[ , colSums(is.na(obs)) == 0]               # Remove columns with at least one NA

# Save it
# saveRDS(obs, file.path(".", "processed", "obs_notright.rds"))
# obs_nr <- readRDS(file.path(".", "processed", "obs_notright.rds"))

#saveRDS(obs, file.path(".", "processed", "obs.rds"))
```

List of problems:
- the s_sf state, storage in the surface, should be less than 1 everywhere that is not flowing, or stream network reaches.

# RUN MODEL
Fixed running model I think!
```{r}
rm(list=ls())

## Load model and data
mdl <- readRDS(file.path(".", "dynaGIS", "atb_20.rds"))
mdl2 <- readRDS(file.path(".", "dynaGIS", "VERYGOODMODEL.rds"))
#obs <- readRDS(file.path(".", "processed", "obs.rds"))
hru <- readRDS(file.path(".", "dynaGIS", "VERYGOODMODEL_hru.rds"))
## ── FIX 1 & 2: Convert precip and PET from mm/step to m/step ──────────────
#obs$precip <- obs$precip / 1000
#obs$pet    <- obs$pet    / 1000

fpar <- function(h){ ## h is the hru
    ## test if a channel
    if( !is.na(h$class$startNode) ){
        ## change type
        h$sf$type <- "kin"
        h$sf$param["c_sf"] <- 1
    }
    h
}

fpar <- function(h){ ## h is the hru
    
        h$uz$parameters["t_d"] <- 104.775
        h$sz$parameters["m"] <- 0.00698783
    h
}
hru <- lapply(hru, fpar)

chn_id <- do.call(c, lapply( hru, function(h){ if(!is.na(h$class$startNode)){ h$id } }))

out_hyd <-  data.frame( name = paste0("hyd_",chn_id),
                       id = chn_id,
                       flux = "v_sf_rz", scale =-1/900) # scale is negative since flux usually downwards, 1/900 is 1/timestep

output_defn <- out_hyd[2,]


ctch_mdl <- dynatop$new(hru, map = mdl$map)$add_data(obs)$initialise()$sim(output_defn)


plot(ctch_mdl$get_output(), legend.loc='topright')

plot(merge(ctch_mdl$get_output(), obs$Discharge_m3_s),  main="Modeled discharge",ylab="m^3",
        legend.loc='topright')

1 - sum((obs$Discharge_m3_s-ctch_mdl$get_output())^2,na.rm=TRUE) /sum((obs$Discharge_m3_s - mean(obs$Discharge_m3_s,na.rm=TRUE))^2,na.rm=TRUE)
```


# CLEAN RUN, ITERATE ON AVAILABLE YEARS
```{r}

years <- seq(17, 24, 1)

for(i in 1:length(years)){
start <- paste0("20",years[i],"-05-01")
end <- paste0("20",years[i],"-10-31")
###### Create obs ######
flow_xts <- 
  flow_15min %>%
  filter(DATETIME >= start,
         DATETIME <= end) %>% 
  # Create a new column that groups time into 15-minute buckets
  mutate(time_15 = floor_date(DATETIME, "15 mins")) %>%
  # Group by those buckets
  group_by(time_15) %>%
  # Average the 5-minute instantaneous rates to get the 15-minute rate
  summarize(Discharge_ls = mean(Discharge_ls, na.rm = TRUE)) %>%
  # convert to m3_s, or should it be m3/day?>
  mutate(Discharge_m3_s = Discharge_ls / 1000) %>% 
  as.xts()


precip_xts <- 
  precip %>% 
  filter(DateTime >= start,
         DateTime <= end) %>% 
  select(DateTime, precip) %>% 
  mutate(precip = precip / 1000) %>% # convert to m
  as.xts()


air_formatted <- air %>% 
  filter(DateTime >= start,
         DateTime <= end) %>% 
  select(DateTime, airTemp)

air_formatted$PET_mm_day <- hargreaves_pet(
  air_formatted$DateTime,
  air_formatted$airTemp,
  lat
)

pet_xts <- air_formatted %>% 
  select(DateTime, PET_mm_day) %>% 
  mutate(PET_m_day = PET_mm_day/1000/1000/96) %>% # convert to m, and 15 min
  rename(pet = PET_m_day) %>% 
  select(DateTime, pet) %>% 
  as.xts()


obs <- merge(flow_xts,
             precip_xts,
             pet_xts,
             all=FALSE)

###### Run Model ######
ctch_mdl <- dynatop$new(hru, map = mdl$map)$add_data(obs)$initialise()$sim(output_defn)


datetime <- index(obs)
actual <- obs$Discharge_m3_s
modeled <- ctch_mdl$get_output()

# return nse
nse <- 1 - sum((obs$Discharge_m3_s-ctch_mdl$get_output())^2,na.rm=TRUE) /sum((obs$Discharge_m3_s - mean(obs$Discharge_m3_s,na.rm=TRUE))^2,na.rm=TRUE)

output <- data.frame(
  "datetime" = datetime,
  "actual" = actual,
  "modeled" = modeled,
  "nse" = nse
) %>% tibble()

if (i == 1) alldat <- output
if (i > 1) alldat <- rbind(alldat, output)
}
colnames(alldat) <- c("datetime",
                      "Discharge",
                      "Model",
                      "NSE")
alldat %>% 
  mutate(year = year(datetime)) %>% 
  pivot_longer(-c(datetime, NSE, year)) %>% 
  ggplot(aes(x = datetime, y = value, color = name))+
  geom_line()+
  theme_classic()+
  facet_wrap(~year, scales = "free")+
  scale_color_manual(values = c("black",
                                "salmon"))

alldat %>% 
  mutate(year = year(datetime)) %>% 
  select(year, NSE) %>% 
  unique()
```





```{r identifying channels}

chn_id <- do.call(c, lapply( hru, function(h){ if(!is.na(h$class$startNode)){ h$id } }))

out_hyd <-  data.frame( name = paste0("hyd_",chn_id),
                       id = chn_id,
                       flux = "v_sf_rz", scale =-1/900) # scale is negative since flux usually downwards, 1/900 is 1/timestep

output_defn <- out_hyd[2,]
## function to compute the relevent part so the output definition for each potential source HRU
fout <- function(h){
    ## test if it is a channel HRU - in which case return NULL
    if( !is.na(h$class$startNode) ){ return(NULL) }
        
    ## index of flow directiuons going to other channels
    idx <- h$sf_flow_direction$id %in% chn_id
    if(!any(idx)){ return(NULL) } ## quit if no links!
    ## create partial output defn
    data.frame(
        name = paste0("hyd_",h$sf_flow_direction$id[idx]), ## which inflow does it alter
        id = h$id, ## flux comes from this HRU
        flux = "q_sf", # surface flux out of this HRU is input to the next
        scale = h$sf_flow_direction$fraction[idx] # scale is fraction of flow (since taking it away)
    )
}
## row bind to give a data.frame
tmp <- do.call( rbind, lapply( hru, fout) )

## row bind onto the intial data frame
out_hyd <- rbind( out_hyd, tmp )

## sort this, not needed but easier to look at
out_hyd <- out_hyd[ order(out_hyd$name), ]

ctch_mdl <- dynatop$new(hru)$add_data(obs)$initialise()$sim(out_hyd)
plot( ctch_mdl$get_output(), legend.loc='topright')
```

```{r}
## extract the current states as a data.frame
current_state <- ctch_mdl$get_states()
## rows correspond to the Hill-slope HRU ID, and the storage's in each of the four zones
head(current_state)

ctch_mdl$plot_state("s_sf")
ctch_mdl$plot_state("s_rz")
ctch_mdl$plot_state("s_uz")
ctch_mdl$plot_state("s_sz")
```

